\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage[hidelinks]{hyperref}
\usepackage{enumitem}

\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\theoremstyle{definition}
\newtheorem{definition}{Definition}

\newcommand{\E}{\mathbb E}
\newcommand{\Prb}{\mathbb P}
\newcommand{\Bare}{\operatorname{Bare}}

\title{Review-Focused Replacement Text for Sections 7--9}
\author{Companion revision for the Erd\H{o}s Problem 625 candidate manuscript}
\date{23 July 2026}

\begin{document}
\maketitle

\paragraph{Status.}
This file supplies drop-in replacement passages for the most concentrated
parts of the canonical manuscript.  It does not change the theorem, constant,
four class sizes, or proof architecture.  Its purpose is to make the
quantifiers, finite decompositions, and multiplicity accounting explicit
enough for independent review and later formalization.  Equation numbers refer
to the canonical manuscript.

\section*{Global notation and uniformity edits}

Use
\[
 q=\ln2,\qquad N=\ln n,\qquad w=\ln N.
\]
Rename the affine coefficient in equation (3.12) from $B_n$ to
$b_n^{\mathrm{aff}}$.  Reserve
\[
 \mathcal B_n:=\frac{n}{N^4}
\]
for the amplification scale currently called $B_n$ in equation (10.10).

Unless a statement says otherwise, every asymptotic estimate below is uniform
over the complete phase interval $0\le\delta<1$, the exact tangent-rounded
four-size profile, every feasible subprofile or canonical skeleton in the
displayed finite sum, and every residual degree list satisfying the displayed
cap and total-mass hypotheses.  Constants may depend on fixed numerical
corridor widths, but not on the phase, profile coordinates, skeleton, or
residual table.

\section*{Replacement A: the central rate in Lemma 7.1}

\begin{lemma}[Uniform central-rate gap]
Let $p=(p_i)_{i=2}^5$ be a probability vector with
\[
 \sum_{i=2}^5 i p_i=T,
 \qquad
 \frac2q\le T\le1+\frac2q.
\]
Let $0\le z_i\le p_i$, and put
\[
 R=\sum_{i=2}^5 z_i,
 \qquad
 Y=1-R,
 \qquad
 I_r=\sum_{i=2}^5 i z_i.
\]
For
\[
 \Phi_T(z)=R\ln R+\frac q2(I_r-TR),
\]
with $0\ln0=0$, one has
\[
 \Phi_T(z)\le-\frac{Y}{5000}
 \qquad\text{whenever}\qquad
 \frac1{64}\le R\le1.
 \tag{7.21a}
\]
\end{lemma}

\begin{proof}
Because every residual deficit lies in $\{2,3,4,5\}$,
\[
 I_r-TR=\sum_i(i-T)z_i\le(5-T)R.
 \tag{7.21b}
\]
Since $y_i=p_i-z_i$, the mean identity for $p$ also gives
\[
 I_r-TR=\sum_i(T-i)y_i\le(T-2)Y.
 \tag{7.21c}
\]
First suppose $1/64\le R\le47/100$.  Since $T\ge2/q$,
\[
 \Phi_T(z)
 \le R\left\{\ln R+\frac q2\left(5-\frac2q\right)\right\}
 =R\{\ln R+5q/2-1\}.
 \tag{7.21d}
\]
The function
\[
 f(R)=R\{\ln R+5q/2-1\}+\frac{1-R}{5000}
\]
is convex on $(0,\infty)$.  Hence its maximum on
$[1/64,47/100]$ occurs at an endpoint.  Using $q<0.6932$,
\[
 f(1/64)<-0.053,
 \qquad
 f(47/100)<-0.010,
\]
so $f(R)<0$ throughout the interval.

Now suppose $47/100\le R\le1$.  Since $T\le1+2/q$, equation (7.21c)
gives
\[
 \Phi_T(z)
 \le R\ln R+(1-q/2)(1-R).
 \tag{7.21e}
\]
Set
\[
 h(R)=R\ln R+(1-q/2+1/200)(1-R).
\]
This function is convex, $h(1)=0$, and direct evaluation gives
$h(47/100)<-0.0058$.  Convexity therefore places $h$ below the chord joining
these endpoint values, so $h(R)\le0$ on $[47/100,1]$.  Thus
\[
 \Phi_T(z)\le-\frac{1-R}{200}
 \le-\frac{1-R}{5000}.
\]
\end{proof}

\paragraph{Completion of the central range.}
If
\[
 m>\eta n,
 \qquad
 n-m>n/32,
\]
then equation (7.13) gives $R\ge1/64$ for all sufficiently large $n$, and
the selected mass satisfies $Y\ge\eta/2$.  Equation (7.20) and the lemma give
\[
 \ln D(\ell)
 \le-\frac{k_{co}\alpha Y}{5000}
      +Ck_{co}Y\ln(e/Y)+CN.
 \tag{7.24a}
\]
Here $Y\ge w/(64N)$, $\alpha=(2/q+o(1))N$, and
$k_{co}=\Theta(n/N)$.  Consequently
\[
 \frac{\ln(e/Y)}{\alpha}=o(1),
 \qquad
 \frac{N}{k_{co}\alpha Y}=o(1),
\]
uniformly.  After increasing the eventual threshold for $n$, the error terms
are at most half of the leading negative term.  Hence, for an absolute $c>0$,
\[
 D(\ell)\le \exp(-c k_{co}w).
 \tag{7.25}
\]
The at most $(k_{co}+1)^4$ central vectors therefore contribute $o(1)$.

\section*{Replacement B: exact canonical exposure at the start of Section 8}

\begin{proposition}[Canonical high-cell exposure identity]
Fix row degrees $(s_a)_a$ and column degrees $(t_b)_b$, each summing to $n$,
and let $r=(r_{ab})$ be a feasible overlap table.  Put
\[
 U=\alpha-2,
 \qquad
 R_0=\lfloor U/2\rfloor,
 \qquad
 M(r)=\{(a,b):r_{ab}>R_0\}.
\]
Then:
\begin{enumerate}[label=\arabic*.]
\item $M(r)$ is a partial matching.
\item For $e=(a,b)\in M(r)$, set $j_e=r_{ab}$ and
$J=\sum_{e\in M(r)}j_e$.  The residual degrees are
\[
 s_a'=s_a-\sum_{b:(a,b)\in M(r)}j_{ab},
 \qquad
 t_b'=t_b-\sum_{a:(a,b)\in M(r)}j_{ab}.
\]
\item The residual table
\[
 r'_{ab}=
 \begin{cases}
 0,&(a,b)\in M(r),\\
 r_{ab},&(a,b)\notin M(r)
 \end{cases}
\]
has those residual margins, vanishes on $M(r)$, and is at most $R_0$ off the
matching.
\item The matching, demands, selected labelled stub pairs, and capped residual
matching reconstruct the original matching uniquely.
\item The exposure incidence times the residual contingency law equals the
original table law exactly.
\end{enumerate}
\end{proposition}

\begin{proof}
The matching property follows because every row and column degree is at most
$U$, while two cells larger than $U/2$ in one row or column would use more than
$U$ stubs.  The margin and cap statements follow directly from the definition.
Reconstruction is by adjoining the selected labelled pairs to the residual
matching; canonical extraction recovers the same data.

Write
\[
 d_a=\sum_{b:(a,b)\in M}j_{ab},
 \qquad
 d_b'=\sum_{a:(a,b)\in M}j_{ab}.
\]
The exposure incidence is
\[
 \pi(M,j)=
 \frac{
   \prod_a(s_a)_{d_a}
   \prod_b(t_b)_{d_b'}
 }{
   (n)_J\prod_{e\in M}j_e!
 }.
 \tag{8.3a}
\]
The residual table mass is
\[
 p_{res}(r')=
 \frac{
   \prod_a(s_a-d_a)!
   \prod_b(t_b-d_b')!
 }{
   (n-J)!\prod_{a,b}r'_{ab}!
 }.
 \tag{8.3b}
\]
Using $(s)_d=s!/(s-d)!$ and $(n)_J=n!/(n-J)!$ gives
\[
\begin{split}
 \pi(M,j)p_{res}(r')
 &=\frac{\prod_a s_a!\prod_b t_b!}
         {n!\prod_{e\in M}j_e!\prod_{a,b}r'_{ab}!}\\
 &=\frac{\prod_a s_a!\prod_b t_b!}
         {n!\prod_{a,b}r_{ab}!}
 =p(r).
\end{split}
\tag{8.3c}
\]
Thus every table appears once, with no hidden multiplicity or proportionality
constant.
\end{proof}

\begin{definition}[Bare canonical-skeleton weight]
For a feasible canonical skeleton $(M,j)$, set
\[
 \Bare(M,j)=\pi(M,j)\prod_{e\in M}g(j_e).
 \tag{8.3d}
\]
The residual cap and no-return event remain in the residual law, while the
capped residual local/cycle factor is postponed to Section 9.
\end{definition}

When an unfinished high-cell completion is bounded above by the full
residual integrand, the latter is only a larger nonnegative numerical
majorant.  It does not redefine $\Bare(M,j)$, and the Section 9 residual factor
is still applied exactly once.

\section*{Replacement C: split form of Lemma 8.3}

\begin{lemma}[One-cell near-containment sum]
Let the smaller slot have size $m$, the larger size $m+d$, where
$0\le d\le3$.  Replacing endpoint multiplicity $m$ by $m-e$ has exact ratio
\[
 R_{m,d}(e)=
 \frac{\binom me}{(d+1)\cdots(d+e)}
 2^{-em+e(e+1)/2}.
 \tag{8.21}
\]
Uniformly in the four cell types,
\[
 \sum_{1\le e<m/4}n^eR_{m,d}(e)=O(N^3/n).
 \tag{8.25}
\]
\end{lemma}

\begin{proof}
The consecutive ratio is
\[
 \rho_e=
 \frac{n(m-e)2^{-m+e+1}}{(e+1)(d+e+1)}.
 \tag{8.23}
\]
For real $e$,
\[
 \frac{d^2}{de^2}\ln\rho_e
 =-\frac1{(m-e)^2}
  +\frac1{(e+1)^2}
  +\frac1{(d+e+1)^2}>0
 \tag{8.24a}
\]
on $0\le e\le m/4$ for all sufficiently large $m$.  Thus the ratios are
log-convex and attain their maximum at an endpoint.  At $e=0$ the ratio is
$O(N^3/n)$, and at $e=\lfloor m/4\rfloor$ it is $n^{-1/2+o(1)}$.  Both are
eventually below $1/2$, uniformly, so the series is geometrically decreasing.
\end{proof}

\begin{lemma}[Global product over near-containment decorations]
Fix a typed full-containment table $L$ and temporarily distinguish every
endpoint cell occurrence.  Let $\mathcal S(L)$ be the family obtained by
choosing for each occurrence either deficit $0$ or $1\le e<m/4$.  If $w(S)$ is
the exact incidence and local high-cell reward, then
\[
 \sum_{S\in\mathcal S(L)}w(S)
 \le W(L)\prod_{c\in L}
 \left(1+\sum_{1\le e<m_c/4}n^eR_{m_c,d_c}(e)\right).
 \tag{8.25a}
\]
Forgetting the temporary labels gives exactly the typed multinomial
multiplicity.  Since a high skeleton has at most $k_{co}$ cells,
\[
 \sum_{S\in\mathcal S(L)}w(S)
 \le W(L)(1+O(N^3/n))^{k_{co}}
 =W(L)\exp(O(N^2)).
 \tag{8.26}
\]
\end{lemma}

\begin{lemma}[Middle strip with large residual mass]
Fix a near-containment skeleton $S$, and let $m_0$ be the remaining stub mass.
On the event $\mathcal N(S)$ that no additional near-containment cell occurs,
every unexposed high cell satisfies
\[
 R_0<j\le3a/4+O(1),
 \qquad a=\alpha-2.
\]
If $m_0\ge n/N^6$, then
\[
 E_{mid}(S)\le\exp(\Xi_4),
 \tag{8.26a}
\]
where
\[
 \Xi_4\le k_{co}^2
 \sum_{a/2<j\le3a/4+O(1)}
 g(j)\frac{(ea^2/m_0)^j}{j!}.
 \tag{8.27}
\]
There is an absolute $c_0>0$ such that, eventually and uniformly,
\[
 \log_2\left[
   k_{co}^2g(j)\frac{(ea^2/m_0)^j}{j!}
 \right]
 \le-c_0(\log_2n)^2.
 \tag{8.28a}
\]
Thus $\Xi_4=2^{-\Omega((\log n)^2)}$.
\end{lemma}

\begin{proof}
Expand over distinct residual cells and threshold demands before dropping any
constraints.  Lemma 6.2 gives (8.26a)--(8.27).  Put $L_2=\log_2n$ and
$j=xL_2$.  Then
\[
 1+o(1)\le x\le3/2+o(1),
 \qquad
 \log_2m_0\ge L_2-6\log_2N.
\]
Using $g(j)\le2^{j^2/2}$ and $k_{co}\le n$,
\[
 \log_2\left[
   k_{co}^2g(j)\frac{(ea^2/m_0)^j}{j!}
 \right]
 \le2L_2+
 \left(\frac{x^2}{2}-x\right)L_2^2
 +O(L_2\log L_2).
\]
On $[1,3/2]$, $x^2/2-x\le-3/8$.  Taking $c_0=1/4$, the lower-order term is
smaller than $L_2^2/8$ for all sufficiently large $n$, uniformly over the floor
perturbations and cell types.
\end{proof}

\begin{lemma}[Small residual completion]
If $m_0<n/N^6$, then pointwise
\[
 \beta(S\cup H_{res})\le|E(H_{res})|\le m_0/2,
 \qquad
 \sum_e\binom{r_e}{2}\le(a-1)m_0/2.
 \tag{8.29}
\]
Hence every nonnegative completion weight needed to dominate the unfinished
high-cell sum is at most
\[
 \exp(Cam_0)\le\exp(Cn/N^5).
 \tag{8.29a}
\]
\end{lemma}

This may use the full residual cycle factor as a pointwise majorant.  It is
only a numerical upper bound for the bare completion sum and does not alter the
exact factorization in Section 9.

\begin{proposition}[Sum of all bare canonical skeletons]
The sum over every feasible canonical matching, allowed high multiplicity, and
induced nonnegative residual degree list satisfies
\[
 \sum_{(M,j)}\Bare(M,j)
 \le e^{O(N^2)}
 \left(\sum_LW(L)\right)
 \max\{e^{\Xi_4},e^{Cn/N^5}\}.
 \tag{8.29b}
\]
Consequently
\[
 \sum_{(M,j)}\Bare(M,j)
 \le\exp\{O(\sqrt{nN})+O(N^2)+O(n/N^5)\}
 =\exp\{o(n/N^4)\}.
 \tag{8.30}
\]
\end{proposition}

Endpoint and near-endpoint multiplicities are generated from a typed endpoint
table, and every remaining high multiplicity is generated in the middle-strip
expansion under the no-further-near-cell event.  Hence every multiplicity is
included exactly once.

\section*{Replacement D: finite cycle expansion in Lemma 9.1}

\begin{lemma}[Even subgraphs to a simple-cycle product]
Let $G$ be a finite graph with nonnegative edge weights $(w_e)$.  For every
even edge set $F$, fix deterministically an edge-disjoint decomposition
$\mathcal D(F)$ into simple cycles.  Then
\[
 \sum_{F\text{ even}}\prod_{e\in F}w_e
 \le
 \prod_{C\text{ simple cycle}}
 \left(1+\prod_{e\in C}w_e\right)
 \le
 \exp\left\{\sum_{C\text{ simple cycle}}\prod_{e\in C}w_e\right\}.
 \tag{9.15a}
\]
\end{lemma}

\begin{proof}
The cycles in $\mathcal D(F)$ are edge-disjoint and therefore distinct, and
\[
 \prod_{e\in F}w_e
 =\prod_{C\in\mathcal D(F)}\prod_{e\in C}w_e.
\]
The union of the cycles recovers $F$, so the map to the resulting cycle set is
injective.  Dropping edge-disjointness enlarges the image to all subsets of the
finite simple-cycle set.  Summing those subsets gives the first product, and
$1+x\le e^x$ gives the second inequality.
\end{proof}

Apply the lemma to $G=M\cup R$, with matching-edge weight one and residual-edge
weight $q_e$.

\begin{lemma}[Cycles disjoint from the high matching]
Let $Q$ be the symmetric weighted adjacency kernel of the residual bipartite
graph.  If every row and column sum is at most $\tau<1$, then
\[
 \sum_{C:C\cap M=\varnothing}\prod_{e\in C}q_e
 \le\frac{n\tau^4}{1-\tau^2}.
 \tag{9.16}
\]
\end{lemma}

\begin{proof}
A residual-only bipartite cycle has length $2s\ge4$.  Mark one row vertex,
forget simplicity, and forget the closing constraint.  For each marked start,
the total mass of length-$2s$ walks is at most $\tau^{2s}$.  There are at most
$n$ marked starts, so summing $s\ge2$ gives the result.
\end{proof}

\begin{lemma}[Cycles meeting the high matching]
Let $M$ be a bipartite matching of size $h$.  Let $Q$ be a nonnegative symmetric
residual kernel, zero on $M$, with row-sum norm at most $\tau<1/3$.  Put
\[
 P=Q+Q^2+Q^3+\cdots,
 \qquad
 b=\|P\|_{\infty}\le\frac{\tau}{1-\tau}.
 \tag{9.17a}
\]
Then
\[
 \sum_{C:C\cap M\ne\varnothing}\prod_{e\in C\setminus M}q_e
 \le2h\sum_{r\ge1}b^r
 =\frac{2hb}{1-b}
 \le Ch\tau.
 \tag{9.18a}
\]
\end{lemma}

\begin{proof}
Fix a total order on the oriented matching edges.  For every simple cycle
meeting $M$, mark the least oriented matching edge on the cycle.  This costs at
most $2h$ choices and removes rotation and orientation ambiguity.

If the cycle uses $r\ge1$ matching edges, cutting at them leaves $r$ nonempty
residual paths.  Between consecutive residual paths, the next matching edge is
determined by the current endpoint because $M$ is a matching.  Matching
traversal is therefore a partial permutation of row-sum norm one and introduces
no new factor $h$.

After dropping simplicity, vertex-disjointness, and the final closing
constraint, each residual path contributes total mass at most $b$.  Relaxed
codes with exactly $r$ matching edges therefore have total mass at most
$2hb^r$.  Summing over $r$ proves the claim.
\end{proof}

\paragraph{Completion of the large-residual branch.}
Equations (9.12)--(9.14) give
\[
 \Lambda_0\le C U^4/m_0,
 \qquad
 \tau\le C U^3/m_0.
\]
For $m_0\ge n/N^6$, one has $\tau<1/3$ eventually, uniformly over every
feasible canonical skeleton.  The preceding lemmas and $h<2n/U$ yield
\[
 \mathcal A(M,j)
 \le
 \exp\left[
 C\left\{
   \frac{U^4}{m_0}+n\tau^4+h\tau
 \right\}
 \right]
 \le\exp(C'N^8).
 \tag{9.19}
\]
This is a one-sided bound.  No lower estimate for $\mathcal A(M,j)$ is asserted
or used.

\section*{Integration checklist}

\begin{enumerate}
\item Preserve equations (7.8)--(7.21), insert the central-rate lemma, and
replace the paragraph leading to (7.25).
\item Insert the canonical exposure proposition and bare-weight definition
after (8.3).
\item Replace the current Lemma 8.3 by the split lemmas and final proposition.
\item Insert the three cycle lemmas between (9.14) and the large-residual
conclusion.
\item Rename the two unrelated uses of $B_n$.
\item Regenerate the TeX and PDFs from the canonical source.
\item Rerun display-tag, citation, finite-diagnostic, and formalization-status
checks.
\item Retain the status ``candidate solution; not externally or fully formally
verified'' until a stronger review is complete.
\end{enumerate}

\end{document}
