\documentclass[11pt,reqno]{amsart}

\usepackage[letterpaper,margin=1.12in]{geometry}
\usepackage[T1]{fontenc}
\usepackage{newtxtext,newtxmath}
\usepackage{microtype}
\usepackage{amsmath,mathtools}
\usepackage[authoryear,round]{natbib}
\usepackage{enumitem}
\usepackage{needspace}
\usepackage{xurl}
\usepackage{xcolor}
\usepackage[hidelinks]{hyperref}
\usepackage{bookmark}

\setcounter{secnumdepth}{1}
\setlength{\parindent}{1.45em}
\setlength{\parskip}{0pt}
\setlist[itemize]{leftmargin=2.05em,itemsep=0.12em,topsep=0.32em}
\setlist[enumerate]{leftmargin=2.15em,itemsep=0.12em,topsep=0.32em}
\setlength{\emergencystretch}{3em}
\allowdisplaybreaks
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\newcommand{\Prob}[1]{\mathbb{P}\!\left(#1\right)}
\newcommand{\Exp}[1]{\mathbb{E}\!\left[#1\right]}
\newcommand{\Expres}[1]{\mathbb{E}_{\mathrm{res}}\!\left[#1\right]}
\newcommand{\whp}{with high probability}
\newcommand{\gnp}{G(n,1/2)}
\newcommand{\gap}{\chi(G_n)-\zeta(G_n)}
\newcommand{\displayheading}[1]{%
  \par\addvspace{0.68\baselineskip}%
  \noindent{\scshape #1.}\par\nobreak\smallskip\noindent}

\definecolor{statementblue}{RGB}{25,55,105}
\definecolor{statementgray}{RGB}{80,80,80}

\newenvironment{resultbox}[1]
  {\Needspace{9\baselineskip}\par\addvspace{0.85\baselineskip}%
   \noindent{\color{statementblue}\rule{\linewidth}{0.55pt}}\par\nobreak
   \noindent{\color{statementblue}\textbf{#1.}}\enspace\itshape\ignorespaces}
  {\par\nobreak\noindent{\color{statementblue}\rule{\linewidth}{0.35pt}}%
   \par\addvspace{0.35\baselineskip}}
\newenvironment{lemmabox}[1]
  {\Needspace{12\baselineskip}\par\addvspace{0.72\baselineskip}%
   \noindent{\color{statementgray}\rule{\linewidth}{0.35pt}}\par\nobreak
   \noindent{\color{statementgray}\textbf{#1.}}\enspace\itshape\ignorespaces}
  {\par\nobreak\noindent{\color{statementgray}\rule{\linewidth}{0.25pt}}%
   \par\addvspace{0.30\baselineskip}}
\newenvironment{propositionbox}[1]
  {\Needspace{8\baselineskip}\par\addvspace{0.72\baselineskip}%
   \noindent{\color{statementgray}\rule{\linewidth}{0.35pt}}\par\nobreak
   \noindent{\color{statementgray}\textbf{#1.}}\enspace\itshape\ignorespaces}
  {\par\nobreak\noindent{\color{statementgray}\rule{\linewidth}{0.25pt}}%
   \par\addvspace{0.30\baselineskip}}

\hypersetup{
  pdftitle={A Polynomial-Scale Gap Between the Chromatic and Cochromatic Numbers of a Random Graph},
  pdfauthor={Samuil Petkov},
  pdfcreator={LaTeX}}

\title[Chromatic and cochromatic numbers]{A Polynomial-Scale Gap Between the
Chromatic and Cochromatic Numbers of a Random Graph}
\author[Samuil Petkov]{Samuil Petkov}
\address{\'Ecole normale sup\'erieure, Universit\'e PSL, Paris, France}
\email{samuil.petkov@ens.psl.eu}
\date{July 12, 2026}
\subjclass[2020]{Primary 05C80; Secondary 05C15, 60C05}
\keywords{random graph, chromatic number, cochromatic number, second moment method}

\begin{document}

\begin{abstract}

For a graph \(G\), let \(\chi(G)\) be its chromatic number and
let \(\zeta(G)\) be the least number of parts in a vertex partition
in which every part induces either an empty graph or a complete graph.
We prove that, for \(G_n\sim G(n,1/2)\),

\[
 \Prob{\chi(G_n)-\zeta(G_n)
 \ge \frac{(\ln 2)^2}{32}\ln\!\left(\frac{200}{153}\right)
       \frac{n}{(\ln n)^3}}\longrightarrow1.             \tag{0.1}
\]

The delicate point is the rounding phase of the independence-number
threshold. We compare ordinary colorings with signed cocolorings whose
classes have four consecutive sizes and may be declared either independent
or complete. The sign choice creates an exquisitely phase-sensitive but
uniform first-moment advantage, lowering the relevant root by order
\(n/(\ln n)^3\). To show that this lower root is attained, we derive an exact
overlap identity and express its sign contribution through even subgraphs of
a bipartite configuration model. The resulting second-moment estimate first
produces a possibly rare cocoloring; a vertex-exposure argument then promotes
that seed to a high-probability event without changing the leading root
separation.

\end{abstract}

\maketitle

\section*{Introduction}\label{introduction}

Let \(G_n\sim G(n,1/2)\) be the random graph on \([n]\), with each edge
present independently with probability \(1/2\).  The chromatic number
\(\chi(G)\) is the least number of independent sets in a partition of
\(V(G)\).  The cochromatic number \(\zeta(G)\) is the least number of parts
in a partition in which every part induces either an empty graph or a
complete graph.  We say that an event holds with high probability if its
probability tends to one as \(n\to\infty\).

Erd\H{o}s and Gimbel asked whether
\(\chi(G_n)-\zeta(G_n)\to\infty\) with high probability
\citep[p.~263]{erdos-gimbel-1993}.  The question was later restated by
\citet[Section~7.4, p.~100]{gimbel-2016} and is catalogued as Erd\H{o}s
Problem~625 \citep{bloom-erdos625}.  We prove the following quantitative
form.

\begin{resultbox}{Theorem 1}
For \(G_n\sim G(n,1/2)\),
\[
 \Prob{\chi(G_n)-\zeta(G_n)
 \ge \frac{(\ln 2)^2}{32}\ln\!\left(\frac{200}{153}\right)
       \frac{n}{(\ln n)^3}}\longrightarrow 1.
 \]
 In particular, \(\chi(G_n)-\zeta(G_n)\) diverges with high probability
 along the full sequence of integers \(n\).
\end{resultbox}

The full-sequence quantifier is essential. Near an independence-number jump,
a rounding error of one vertex changes the available profile, so an estimate
proved only on a dense set of phases does not automatically extend to every
integer \(n\). The argument below is therefore uniform in the complete phase
parameter \(\delta\in[0,1)\), including sequences approaching either side of
each jump.

\displayheading{Background and relation to earlier work}
The chromatic number of dense random graphs has been studied since the
1970s.  \citet[Corollary~3]{grimmett-mcdiarmid-1975} established the correct first-order
lower scale and formulated the corresponding upper-bound problem.
\citet{matula-1987} developed an expose-and-merge approach, and
\citet[Theorem~4]{bollobas-1988} proved the first-order asymptotic.
Subsequent refinements include \citet{mcdiarmid-1990},
\citet[Theorem~1.2]{panagiotou-steger-2009}, and the sharp coloring-rate
bounds of \citet[Theorem~1]{heckel-2018}; see also
\citet[Theorem~7.14 and Remark~7.15]{janson-luczak-rucinski-2000}.

For the chromatic--cochromatic gap, \citet[Theorem~1]{heckel-2024-question}
and, independently, \citet[Theorem~1.7]{steiner-2024} gave the first quantitative
evidence toward divergence.  Heckel proved that if a deterministic integer
sequence \(g(n)\) satisfies
\(\Prob{\chi(G_n)-\zeta(G_n)\le g(n)}>0.999\), then along an unbounded
sequence \(n^*\),
\[
 g(n^*)>c\frac{\sqrt{n^*}\,\ln\ln n^*}{(\ln n^*)^3}.
\]
This obstructs uniformly small high-probability upper bounds, but is not an
all-\(n\) high-probability lower bound.  Heckel also conjectured the scale
\(\Theta(n/(\ln n)^3)\) \citep[Conjecture~4]{heckel-2024-question}.

The subsequent theorem of \citet[Theorem~1]{heckel-2025-difference} proved
that, for every fixed \(\varepsilon>0\),
\[
 n^{0.05+\varepsilon}\le\mu_\alpha\le n^{1-\varepsilon}
 \quad\Longrightarrow\quad
 \chi(G_n)-\zeta(G_n)\ge n^{1-\varepsilon}
 \quad\text{with high probability}.
\]
Thus the divergence question was settled on a phase-dependent set comprising
roughly \(95\%\) of the integers; see
\citet[Section~2]{heckel-2025-difference}.

Two ingredients of that work are strategic antecedents here.  First, the
one-partition factor \(2^k\) and the need to count sign compatibility through
the overlap matrix appear in
\citet[Proposition~6]{heckel-2025-difference}.  Section~5 rederives the first
identity, while Section~6 replaces the pairwise upper bound by an exact
component formula.  Second, the rare-seed and concentration-amplification
strategy follows \citet[Proposition~5 and the proof of
Theorem~1]{heckel-2025-difference}.  Sections~6--9 prove the signed
second-moment estimate needed here, and Section~10 proves the quantitative
amplifier.  The four-size profile, its phase-uniform entropy certificate,
the treatment of all remaining overlap regimes, the all-\(n\) conclusion,
and the explicit constant in Theorem~1 are the contributions of this paper.

\displayheading{Proof roadmap}
Sections~2--4 locate \(\chi(G_n)\) from below without imposing a hidden
profile restriction. Section~2 resolves the complete independence-number
phase, Section~3 develops the continuous root geometry, and Section~4 turns
that geometry into an unrestricted chromatic lower bound.

Section~5 constructs the four-size signed profile and proves the decisive
first-moment separation of order \(n/(\ln n)^3\). The rest of the paper shows
that this formal root is genuinely populated. Section~6 derives the exact
overlap and sign decomposition; Section~7 controls common whole classes;
Section~8 sums every canonical high-cell skeleton; and Section~9 bounds the
remaining conditional local and cycle attachments. Section~10 then amplifies
the possibly rare second-moment seed to a high-probability cocoloring event.
Section~11 subtracts the two locations. The roadmap records dependencies
only; each asymptotic estimate is proved where it is used.

\displayheading{How to read the proof}
The proof has three layers, which are best kept separate.  The first layer is
the \emph{location calculation}: Sections~2--5 compare the ordinary coloring
root with the signed four-size cocoloring root.  This is where the constant
\((\ln 2)^2\ln(200/153)/32\) enters.  The second layer is the
\emph{second-moment calculation}: Sections~6--9 prove that the signed
first-moment advantage is not destroyed by overlaps.  These sections contain
most of the bookkeeping; their role is to bound one normalized second moment,
not to move either root.  The third layer is the \emph{probability
amplification}: Section~10 turns a possibly rare cocoloring witness into a
high-probability upper bound for \(\zeta(G_n)\), and Section~11 compares it
with the chromatic lower bound.

The main technical difficulty is that the second layer has to count overlaps
without losing the phase-uniform constant from the first layer.  To make the
bookkeeping auditable, the overlap matrix is reduced in stages.  Section~6
gives an exact sign identity.  Section~7 removes common whole classes.
Section~8 handles the high-multiplicity cells of the remaining overlap
matrix.  Section~9 handles the residual local factors and cycle-space
attachments.  Thus Lemmas~8.3 and~9.1 are not independent black boxes: Lemma
8.3 bounds the high skeletons before the residual graph is exposed, while
Lemma~9.1 bounds what remains after one such skeleton has been fixed.

\section{Notation and elementary
facts}\label{notation-and-elementary-facts}

All logarithms are natural unless a base is displayed.

For an integer \(s\), let

\[
 \mu_s(v)=\binom vs2^{-\binom s2},\qquad \mu_s=\mu_s(n). \tag{1.1}
\]

We use the following standard inequalities in their stated forms.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  For integers \(m\ge 1\), \[
   \ln(m!)=m\ln m-m+\frac12\ln(2\pi m)+O(1/m).          \tag{1.2}
  \] In particular, uniformly for \(m\ge 0\), with
  \(0\ln 0=0\), the last two terms may be replaced by
  \(O(\ln(m+1))\).
\item
  If a random variable \(Y\) is a function of \(r\)
  independent blocks and changing one block changes \(Y\) by at
  most one, then \[
   \Prob{\lvert Y-\Exp{Y}\rvert\ge t}
   \le2\exp(-2t^2/r).                                    \tag{1.3}
  \] We will use the corresponding one-sided bounds as well.
\item
  If \(Z\ge 0\) and \(0<\Exp{Z}<\infty\), then \[
   \Prob{Z>0}\ge\frac{\Exp{Z}^2}{\Exp{Z^2}}.             \tag{1.4}
  \]
\item
  If \(X\sim\operatorname{Bin}(m,1/2)\), then \[
   \Prob{X\le m/4}\le e^{-m/16}.                         \tag{1.5}
  \] Indeed, exponential Markov with \(t=\ln 3\) bounds the
  probability by
  \([3^{1/4}(2/3)]^m\le e^{-m/16}\).
\item
  For every nonnegative random variable \(X\) and \(a>0\),
  \(\Prob{X\ge a}\le \Exp{X}/a\).
\end{enumerate}

The first is Stirling's estimate, the second is McDiarmid's bounded-
differences inequality, and the third is the zero-threshold case of
Paley--Zygmund. The fourth is the only binomial-tail estimate used
below; the fifth is Markov's inequality. The bounded-differences
formulation is the one recorded by
\citet[Theorem~3.1]{mcdiarmid-1989}.

\section{The complete independence-number
phase}\label{the-complete-independence-number-phase}

Define

\[
 \alpha_0=2\log_2 n-2\log_2\log_2 n
             +2\log_2(e/2)+1,
 \quad \alpha=\lfloor\alpha_0\rfloor,
 \quad \delta=\alpha_0-\alpha.                           \tag{2.1}
\]

Thus \(0\le\delta<1\). The estimates below are
uniform on this entire interval, including sequences approaching either
endpoint.

\phantomsection
\addcontentsline{toc}{subsection}{Lemma 2.1 (phase expansion)}
\label{lemma-2.1-phase-expansion}
\begin{lemmabox}{Lemma 2.1 (phase expansion)}

There is a bounded continuous function
\(K\colon[0,1]\to\mathbb R\) such that

\[
 \ln\mu_\alpha-K(\delta)
 =\delta {\ln n}+\left(\frac2{\ln 2}-\frac12-\delta\right){\ln\ln n}
   +O((\ln\ln n)^2/{\ln n}).                                  \tag{2.2}
\]

In particular,

\[
 \mu_{\alpha+2}=n^{\delta-2+o(1)}
 \le n^{-1+o(1)}=o(1),                                   \tag{2.3}
\]

and, with \(a=\alpha-2\),

\[
 \mu_a\ge c n^2(\ln n)^{2/{\ln 2}-5/2}                              \tag{2.4}
\]

for an absolute \(c>0\) and all sufficiently large
\(n\).

\end{lemmabox}

\paragraph{Proof.}\label{proof}

Write

\[
 C=1+\ln(\ln 2)-{\ln 2},\qquad S={\ln n}-{\ln\ln n}+C,\qquad b=1-\delta.
\]

Then \(\alpha=2S/{\ln 2}+b\). Since \(\alpha=O({\ln n})\), (1.2) gives

\[
 \ln(n)_\alpha=\alpha {\ln n}+
 \sum_{j=0}^{\alpha-1}\ln(1-j/n)
 =\alpha {\ln n}+O(\alpha^2/n),
\]

and hence

\[
 \ln\mu_\alpha
 =\ln(n)_\alpha-\ln(\alpha!)
   -{\ln 2}\binom{\alpha}{2}.
\]

Applying Stirling's formula to \(\alpha!\) in this identity gives

\[
 \ln\mu_\alpha
 =\alpha\left({\ln n}-\ln\alpha+1-\frac {\ln 2}2(\alpha-1)\right)
  -\frac12\ln(2\pi\alpha)+O(1/{\ln n}).                       \tag{2.5}
\]

Expanding \(\ln \alpha\) at \(2{\ln n}/{\ln 2}\), uniformly in
\(b \in (0,1]\), gives

\[
 {\ln n}-\ln\alpha+1-\frac {\ln 2}2(\alpha-1)
 =\frac{{\ln 2}\delta}{2}
  +\frac{{\ln\ln n}-C-b{\ln 2}/2}{{\ln n}}+O((\ln\ln n)^2/(\ln n)^2).                        \tag{2.6}
\]

Put \(q=\ln 2\) and \(\ell=\ln\ln n\). Since
\(\alpha=2(\ln n-\ell+C)/q+b\), multiplying the right side of
(2.6) by \(\alpha\) gives

\[
\begin{split}
\alpha\left\{\frac{q\delta}{2}
 +\frac{\ell-C-bq/2}{\ln n}
 +O\!\left(\frac{\ell^2}{(\ln n)^2}\right)\right\}
={}&\delta\ln n+\left(\frac2q-\delta\right)\ell\\
&+\delta C+\frac q2\delta(1-\delta)-\frac{2C}{q}
 -(1-\delta)+O\!\left(\frac{\ell^2}{\ln n}\right).
\end{split}
\]

Moreover,

\[
 -\frac12\ln(2\pi\alpha)
 =-\frac12\ell-\frac12\ln(2\pi)-\frac q2
   +\frac12\ln q+O\!\left(\frac{\ell}{\ln n}\right).
\]

Substitution in (2.5) proves (2.2), with

\[
\begin{split}
 K(\delta)={}&\delta C+\frac {\ln 2}2\delta(1-\delta)-\frac{2C}{{\ln 2}}
 -(1-\delta)\\
 &-\frac12\ln(2\pi)-\frac {\ln 2}2+\frac12\ln(\ln 2).
\end{split}                                               \tag{2.7}
\]

The exact adjacent-size ratios are

\[
 \frac{\mu_{s+1}}{\mu_s}
 =\frac{n-s}{s+1}2^{-s},\qquad
 \frac{\mu_{s-1}}{\mu_s}
 =\frac{s}{n-s+1}2^{s-1}.                                \tag{2.8}
\]

At \(s=\alpha+O(1)\), the first is \(\Theta({\ln n}/n)\) and the
second is \(\Theta(n/{\ln n})\), uniformly in \(\delta\).  Thus

\[
 \mu_{\alpha+2}
 =\mu_\alpha\,\Theta\!\left(\frac{(\ln n)^2}{n^2}\right),
 \qquad
 \mu_{\alpha-2}
 =\mu_\alpha\,\Theta\!\left(\frac{n^2}{(\ln n)^2}\right). \tag{2.8a}
\]

Taking logarithms in the first relation of (2.8a) and using (2.2) gives

\[
 \ln\mu_{\alpha+2}
 =(\delta-2)\ln n+O(\ln\ln n),
\]

which proves (2.3), uniformly in the phase. For the lower adjacent size,
\(\delta(\ln n-\ln\ln n)\ge0\), while \(K\) is bounded, so (2.2) gives

\[
 \ln\mu_\alpha
 \ge\left(\frac2{\ln 2}-\frac12\right)\ln\ln n
   +\inf_{\delta\in[0,1]}K(\delta)+o(1).
\]

Hence \(\mu_{\alpha}\ge c_1(\ln n)^{2/{\ln 2}-1/2}\).
Taking logarithms in the second relation of (2.8a), or equivalently
multiplying this lower bound by \(\Theta(n^2/(\ln n)^2)\), proves
(2.4). \(\square\)

Let \(X_{\alpha+2}\) count independent sets of size \(\alpha+2\).
Then \(\Exp{X_{\alpha+2}}=\mu_{\alpha+2}\), and
\(\alpha(G_n)>\alpha+1\) implies \(X_{\alpha+2}\ge1\).
Thus (2.3) and Markov's inequality give

\[
 \Prob{\alpha(G_n)>\alpha+1}=o(1).                       \tag{2.9}
\]

\section{Continuous profile roots}\label{continuous-profile-roots}

For a class size \(u\), put

\[
 d_u=2^{\binom u2}u!.
\]

Write \(u=\alpha-i\), so that \(i\) is the deficit from
\(\alpha\). Let

\[
 S_+=\{-1,0,1,2,\ldots\},\qquad S_4=\{2,3,4,5\}.         \tag{3.1}
\]

At finite \(n\), \(S_+\) is cut off at \(i=\alpha-1\),
since class sizes are positive. For a real class count \(k\),
define

\[
 L_{\mathbf k}=n{\ln n}-n+k-\sum_i k_i\ln(k_id_{\alpha-i}),    \tag{3.2}
\]

and let \(L_S(n,k)\) be its maximum over real
\(k_i\ge 0\) satisfying

\[
 \sum_{i\in S}k_i=k,\qquad
 \sum_{i\in S}(\alpha-i)k_i=n.                           \tag{3.3}
\]

Set

\[
 T=\alpha-\frac nk.                                      \tag{3.4}
\]

The profile optimization has two jobs. It must locate the zero of the
first-moment exponent, and it must compare two different supports at the
same average class size. The affine part of the deficit weight cancels under
the fixed-mean constraint; only the curved part distinguishes the supports.
The following lemma packages both facts, together with the uniform derivative
needed to convert an entropy advantage into a root displacement.

\phantomsection
\addcontentsline{toc}{subsection}{Lemma 3.1 (root phase, derivative, and support comparison)}
\label{lemma-3.1-root-phase-derivative-and-support-comparison}
\begin{lemmabox}{Lemma 3.1 (root phase, derivative, and support comparison)}

\textup{Let}

\[
 s_0=\alpha_0-1-\frac2{\ln 2},\qquad
 T_0=\alpha-s_0=1+\frac2{\ln 2}-\delta.                        \tag{3.5}
\]

For \(S\in\{S_+,S_4\}\), and uniformly for \(0\le c\le {\ln 2}\), the equation
\(L_S(n,k)+ck=0\) has a unique zero \(r_{S,c}\) in the root corridor
described below, and

\[
 \frac n{r_{S,c}}=s_0+O({\ln\ln n}/{\ln n}),\qquad
 T=\alpha-\frac n{r_{S,c}}=T_0+O({\ln\ln n}/{\ln n}).                  \tag{3.6}
\]

For every fixed \(A>0\), uniformly when \(|n/k-s_0|\le A{\ln\ln n}/{\ln n}\),

\[
 \frac{\partial}{\partial k}\{L_S(n,k)+ck\}
 =\frac2{\ln 2}(\ln n)^2+O({\ln n}{\ln\ln n}).                                      \tag{3.7}
\]

For two supports \(R,S\), at the same feasible \(k\),

\[
 \frac{L_R(n,k)-L_S(n,k)}k
 =\mathcal F_{n,R}(T)-\mathcal F_{n,S}(T),               \tag{3.8}
\]

where

\[
 \mathcal F_{n,S}(T)=
 \max_{\substack{\sum p_i=1\\\sum ip_i=T}}
 \left[-\sum_i p_i\ln p_i+\sum_i p_i h_n(i)\right].     \tag{3.8a}
\]

To make the uniformity precise, fix

\[
 K_*=[2/{\ln 2}-1/10,\,1+2/{\ln 2}+1/10]\subset(2,5).               \tag{3.9a}
\]

Then, uniformly for \(T\in K_*\),

\[
 \mathcal F_{n,S}(T)\longrightarrow
 \mathcal F_S(T):=\max_{\substack{\sum p_i=1\\\sum ip_i=T}}
 \left[-\sum p_i\ln p_i-\frac {\ln 2}2\sum i^2p_i\right].  \tag{3.9}
\]

The effective Lagrange tilts are uniformly bounded on \(K_*\). For
\(S_4\), if \(p_{n,i}(T)\) and \(p_i(T)\) denote the finite and limiting
optimizers, then

\[
 \sup_{T\in K_*}\max_{2\le i\le5}|p_{n,i}(T)-p_i(T)|=o(1),
 \qquad
 \inf_{T\in K_*}\min_{2\le i\le5}p_{n,i}(T)>0.          \tag{3.9b}
\]

All convergence and error terms are uniform in \(\delta\).

\end{lemmabox}

\paragraph{Proof.}\label{proof-1}

Put \(s=n/k\) and \(p_i=k_i/k\). Dividing (3.2) by
\(k\) gives

\[
 \frac{L_S(n,n/s)}{n/s}
 =(s-1){\ln n}-s+1+\ln s
 +\max_p\left[-\sum p_i\ln p_i-\sum p_i\ln d_{\alpha-i}\right].
                                                               \tag{3.10}
\]

There is an exact affine-plus-curved decomposition

\[
 -\ln d_{\alpha-i}=A_n+B_ni+h_n(i),                      \tag{3.11}
\]

where

\[
 A_n=-\ln d_\alpha,\qquad
 B_n={\ln 2}\alpha-{\ln 2}/2+\ln\alpha,                              \tag{3.12}
\]

and, for \(i\ge 0\),

\[
 h_n(i)=-\frac {\ln 2}2i^2+\sum_{r=0}^{i-1}\ln(1-r/\alpha).   \tag{3.13}
\]

For \(i=-1\), direct subtraction gives
\(h_n(-1)=-{\ln 2}/2+\ln(\alpha/(\alpha+1))\).
Thus, for fixed \(i\),

\[
 h_n(i)=-{\ln 2}i^2/2+O(i^2/\alpha),                           \tag{3.14}
\]

while (3.13) also supplies the uniform bound \(h_n(i)\le-{\ln 2}i^2/2\).
The target mean \(T_0\) stays in the compact interval

\[
 \frac2{\ln 2}\le T_0\le1+\frac2{\ln 2}.                             \tag{3.15}
\]

To control the optimizing tilts uniformly, define
\[
 Z_{n,S}(\lambda)=\sum_{i\in S}
   \exp\{\lambda i+h_n(i)\},\qquad
 M_{n,S}(\lambda)=\frac{d}{d\lambda}\ln Z_{n,S}(\lambda),
\]
with the finite-\(n\) cutoff understood for \(S_+\). Let \(Z_S\) and
\(M_S\) denote the corresponding limits obtained by replacing
\(h_n(i)\) with \(-(\ln 2)i^2/2\). Since \(K_*\) lies strictly inside
the convex hull of each support, there is a fixed \(\Lambda>0\) such that
\[
 M_S(-\Lambda)<\min K_*<\max K_*<M_S(\Lambda)
 \qquad(S=S_+,S_4).
\]
On \([-\Lambda,\Lambda]\), the weights and their first two moments are
dominated by a constant multiple of
\(\exp(C|i|-(\ln 2)i^2/2)\). Hence \(Z_{n,S}\), \(M_{n,S}\), and
\(M'_{n,S}\) converge uniformly to their limiting counterparts. In
particular, the preceding strict inequalities also hold with \(M_{n,S}\)
for all sufficiently large \(n\).

The derivative \(M'_{n,S}\) is the variance of \(i\) under the tilted
law. The limiting variance is positive and continuous on the compact
interval \([-\Lambda,\Lambda]\), so uniform convergence gives a constant
\(v_*>0\) such that
\[
 M'_{n,S}(\lambda)\ge v_*
 \qquad(|\lambda|\le\Lambda,\ S=S_+,S_4)
\]
for all sufficiently large \(n\). Thus the inverse mean maps are
uniformly Lipschitz on \(K_*\), and their finite-\(n\) tilts converge
uniformly to the limiting tilts. This proves (3.9), and on the finite
support \(S_4\) it also proves (3.9b). The affine terms in (3.11) cancel
under the fixed mean constraint, giving the exact difference identity
(3.8).

Substitute (3.11) into (3.10). If \(T=\alpha-s\), the scalar part at
\(s=s_0\) is, by Stirling,

\[
 s_0\left({\ln n}-\ln\alpha-\frac {\ln 2}2s_0+\frac {\ln 2}2\right)-{\ln n}
 +T_0+1+\ln s_0+\frac {\ln 2}2T_0^2
 -\frac12\ln(2\pi\alpha)+O(1/{\ln n}).                        \tag{3.15a}
\]

Write \(q=\ln 2\) and \(\ell=\ln\ln n\). Then
\[
 \frac{qs_0}{2}=\ln n-\ell+\ln q-q
\]
exactly, while
\[
 \ln\alpha=\ln s_0+\frac{T_0}{s_0}+O((\ln n)^{-2}),
 \qquad
 \ln s_0=\ell+q-\ln q+O(\ell/\ln n).
\]
Consequently,
\[
\begin{split}
s_0\left(\ln n-\ln\alpha-\frac q2s_0+\frac q2\right)
&=s_0\left(\ell-\ln q+\frac{3q}{2}-\ln\alpha\right)\\
&=s_0\left(\ln s_0-\ln\alpha+\frac q2
             +O(\ell/\ln n)\right)\\
&=-T_0+\frac q2s_0+O(\ell)\\
&=-T_0+\ln n-\ell+\ln q-q+O(\ell).
\end{split}
\]
Substituting this identity into (3.15a) cancels the terms
\(\ln n\) and \(T_0\); the remaining terms are
\[
 -\ell+\ln q-q+1+\ln s_0+\frac q2T_0^2
 -\frac12\ln(2\pi\alpha)+O(\ell)=O(\ell),
\]
uniformly in the phase.
The entropy term is \(O(1)\) by the Gaussian tail just proved.
Hence

\[
 L_S(n,n/s_0)/(n/s_0)=O({\ln\ln n}).                              \tag{3.16}
\]

Write \(\Psi_S(s)=L_S(n,n/s)/(n/s)\), and let
\(\lambda_{n,S}(T)\) be the effective tilt of the maximizing profile at
mean \(T=\alpha-s\).  Differentiating (3.10) by the envelope theorem gives

\[
 \Psi'_S(s)
 ={\ln n}-1+s^{-1}-B_n+\lambda_{n,S}(T).
\]

The tilt is uniformly bounded on the fixed target interval, and

\[
 B_n=2{\ln n}-{\ln\ln n}+O(1),                                          \tag{3.17}
\]

therefore (3.17) gives

\[
 \Psi'_S(s)
 =-{\ln n}+O({\ln\ln n})                                                \tag{3.18}
\]

uniformly in a fixed neighborhood of \(s_0\). For
\(\Psi_{S,c}(s)=\Psi_S(s)+c\), choose \(C>0\) so that
\[
 |\Psi_{S,c}(s_0)|\le C\ln\ln n
\]
for both supports and all \(0\le c\le\ln 2\). By (3.18), throughout a
fixed corridor and for sufficiently large \(n\),
\(\Psi'_{S,c}=\Psi'_S\le-\tfrac12\ln n\). If \(A>2C\), integration gives
\[
\begin{aligned}
\Psi_{S,c}\!\left(s_0-A\frac{\ln\ln n}{\ln n}\right)
 &\ge \left(\frac A2-C\right)\ln\ln n>0,\\
\Psi_{S,c}\!\left(s_0+A\frac{\ln\ln n}{\ln n}\right)
 &\le \left(C-\frac A2\right)\ln\ln n<0.
\end{aligned}
\]
Strict monotonicity therefore gives the unique corridor zero and (3.6),
uniformly for \(0\le c\le\ln 2\).

Throughout any fixed corridor \(|s-s_0|\le A{\ln\ln n}/{\ln n}\), the same equations
give \(\Psi_{S,c}(s)=O_A({\ln\ln n})\). Since \(L_S+ck=k\Psi_{S,c}(s)\) and
\(ds/dk=-s/k\),

\[
 \frac{d}{dk}\{L_S+ck\}
 =\Psi_{S,c}(s)-s\Psi'_S(s)
 =\frac2{\ln 2}(\ln n)^2+O_A({\ln n}{\ln\ln n}).                                    \tag{3.19}
\]

This is (3.7). \(\square\)

\section{\texorpdfstring{A valid unrestricted lower location for
\(\chi\)}{A valid unrestricted lower location for chi}}\label{a-valid-unrestricted-lower-location-for-chi}

Let \(r_+(n)=r_{S_+,0}\) be the zero in Lemma 3.1, and put

\[
 k_\chi^-:=\lfloor r_+(n)\rfloor-\lceil {\ln n}\rceil.         \tag{4.1}
\]

For an integer profile
\(\mathbf k=(k_i)_{-1\le i\le\alpha-1}\), let
\(X_{\mathbf k}\) be the number of unordered proper colorings having
exactly \(k_i\) classes of size \(\alpha-i\). Direct enumeration gives

\[
 \Exp{X_{\mathbf k}}
 =\frac{n!}{\prod_i((\alpha-i)!)^{k_i}k_i!}
  2^{-\sum_i k_i\binom{\alpha-i}{2}}.                   \tag{4.2}
\]

For an integer \(k\), let \(\mathcal K_{n,k,\alpha+1}\) be the set of
such profiles satisfying
\(\sum_i k_i=k\) and \(\sum_i(\alpha-i)k_i=n\), and define
\[
 E_{n,k,\alpha+1}
 :=\sum_{\mathbf k\in\mathcal K_{n,k,\alpha+1}}
   \Exp{X_{\mathbf k}}.
\]
Thus \(E_{n,k,\alpha+1}\) is the expected total number of unordered,
exactly \(k\)-part, \((\alpha+1)\)-bounded colorings. There are at most
\((n+1)^{\alpha+1}=\exp(O((\ln n)^2))\)
bounded profiles. Applying (1.2) to (4.2), uniformly even when some
\(k_i=0\), yields

\[
 \ln E_{n,k,\alpha+1}\le L_+(n,k)+O((\ln n)^2).                \tag{4.3}
\]

By Lemma 3.1,

\[
 L_+(n,k_\chi^-)\le-c(\ln n)^3,                                \tag{4.4}
\]

so the expectation in (4.3) is \(o(1)\). On the event (2.9), any
coloring with at most \(k_{\chi}^-\) nonempty parts can be
refined, by splitting parts, to exactly \(k_{\chi}^-\) nonempty
independent parts, all of size at most \(\alpha+1\). Therefore

\[
 \Prob{\chi(G_n)\le k_\chi^-}=o(1).                     \tag{4.5}
\]

This is an unrestricted chromatic lower bound; the size cap has been
removed probabilistically by (2.9).

\section{The four-size signed first-moment
advantage}\label{the-four-size-signed-first-moment-advantage}

A \emph{signed cocoloring} is a partition in which every class is
declared either ``independent'' or ``complete''; it is realized when each
class induces the graph specified by its declaration.  Thus the sign is a
two-way declaration attached to a class, and the signed counts below count
witnesses for ordinary cocolorings rather than a new graph invariant.

There are two logically separate tasks. First, restricting the deficits to
\(S_4=\{2,3,4,5\}\) must cost strictly less than \(\ln 2\) per part. Second,
the \(2^k\) choices of signs must convert that strict inequality into a
macroscopic separation of the two roots. Lemma~5.1 proves the first point;
the derivative estimate from Lemma~3.1 then proves the second.

For \(S\in\{S_+,S_4\}\), introduce the tilted weights, partition
function, and mean map

\[
 w_i(\lambda):=\exp\!\left(\lambda i-\frac{\ln 2}{2}i^2\right),
 \qquad
 Z_S(\lambda):=\sum_{i\in S}w_i(\lambda),
\]

\[
 M_S(\lambda):=
 \frac{\sum_{i\in S}i\,w_i(\lambda)}{Z_S(\lambda)}
 =\frac{d}{d\lambda}\ln Z_S(\lambda).
\]

The Gaussian factor makes both sums finite.  Moreover,
\(M_S'(\lambda)=\operatorname{Var}_{S,\lambda}(i)>0\), so for every
target mean \(T\) in the interior of the convex hull of \(S\) there is a
unique tilt \(\lambda_S(T)\) with \(M_S(\lambda_S(T))=T\).  The optimizer
in (3.9) is therefore

\[
 p_i=\frac{w_i(\lambda_S(T))}{Z_S(\lambda_S(T))}
 =\frac{e^{\lambda_S(T)i-{\ln 2}i^2/2}}
        {\sum_{j\in S}e^{\lambda_S(T)j-{\ln 2}j^2/2}}.       \tag{5.1}
\]

The same calculation gives the dual representation

\[
 \mathcal F_S(T)
 =\inf_{\lambda\in\mathbb R}
   \{\ln Z_S(\lambda)-\lambda T\}
 =\ln Z_S(\lambda_S(T))-\lambda_S(T)T.
\]

Define

\[
 D_4(\delta)=\mathcal F_{S_+}(T_0)
              -\mathcal F_{S_4}(T_0),\qquad
 \gamma_4=\ln\frac{200}{153}.                            \tag{5.2}
\]

\phantomsection
\addcontentsline{toc}{subsection}{Lemma 5.1 (uniform entropy certificate)}
\label{lemma-5.1-uniform-entropy-certificate}
\begin{lemmabox}{Lemma 5.1 (uniform entropy certificate)}

For every \(0\le \delta\le 1\),

\[
 0\le D_4(\delta)<\ln\frac{153}{100},\qquad
 {\ln 2}-D_4(\delta)>\gamma_4.                                 \tag{5.3}
\]

\end{lemmabox}

\paragraph{Proof.}\label{proof-2}

Fix \(\delta\in[0,1]\), put
\(T_0=1+2/{\ln 2}-\delta\), and abbreviate

\[
 \lambda_4:=\lambda_{S_4}(T_0).
\]

We first bracket this tilt.  At \(\lambda=2{\ln 2}\), we make the
reindexing completely explicit.  The old summation index ranges over
\(i\in S_4=\{2,3,4,5\}\).  Define the new index by
\[
 j:=i-2.
\]
Then \(i\mapsto j=i-2\) is a bijection from
\(\{2,3,4,5\}\) onto \(\{0,1,2,3\}\), with inverse
\(i=j+2\).  Thus the lower and upper bounds change from
\(2\le i\le5\) to \(0\le j\le3\), respectively.  Substituting
\(i=j+2\) in the weight gives

\[
\begin{aligned}
 w_{j+2}(2{\ln 2})
 &=\exp\!\left(2{\ln 2}(j+2)
       -\frac{{\ln 2}}2(j+2)^2\right)\\
 &=\exp\!\left(2{\ln 2}-\frac{{\ln 2}}2j^2\right)
 =4\,2^{-j^2/2}.
\end{aligned}
\]

Consequently, reindexing both sums in the definition of the mean---and
only reindexing them, without discarding any term---gives

\[
\begin{aligned}
 Z_{S_4}(2{\ln 2})
 &=\sum_{i=2}^{5}w_i(2{\ln 2})
   =4\sum_{j=0}^{3}2^{-j^2/2},\\
 \sum_{i=2}^{5}i\,w_i(2{\ln 2})
 &=4\sum_{j=0}^{3}(j+2)2^{-j^2/2}\\
 &=8\sum_{j=0}^{3}2^{-j^2/2}
   +4\sum_{j=0}^{3}j2^{-j^2/2}.
\end{aligned}
\]

Dividing the second equality by the first explains the additive
constant \(2\) in the following formula:

\[
 \begin{aligned}
 M_{S_4}(2{\ln 2})
 &=2+\frac{\sum_{j=0}^{3}j2^{-j^2/2}}
          {\sum_{j=0}^{3}2^{-j^2/2}}\\
 &<2+\frac{\sum_{j\ge0}j2^{-j^2/2}}
          {\sum_{j\ge0}2^{-j^2/2}}
 <2+\frac45<\frac2{\ln 2}\le T_0.
 \end{aligned}
 \tag{5.4}
\]

For completeness, the first strict inequality in (5.4) has the following
direct verification.  Put
\[
 A=\sum_{j=0}^{3}2^{-j^2/2},\quad
 B=\sum_{j=0}^{3}j2^{-j^2/2},\quad
 C=\sum_{j=4}^{\infty}2^{-j^2/2},\quad
 D=\sum_{j=4}^{\infty}j2^{-j^2/2}.
\]
Here \(B/A\le3\), while \(D/C\ge4\).  Since \(A,C>0\),
\[
 \frac{B+D}{A+C}-\frac BA
 =\frac{AD-BC}{A(A+C)}>0.
\]
Thus adjoining the terms whose new indices satisfy \(j\ge4\) strictly
increases the weighted mean. For the next inequality, put
\(a_j=j2^{-j^2/2}\). For \(j\ge4\),
\[
 \frac{a_{j+1}}{a_j}
 =\frac{j+1}{j}\,2^{-(2j+1)/2}\le\frac1{16},
\]
and \(a_4=1/64\). Hence
\[
 \sum_{j\ge4}j2^{-j^2/2}
 \le\frac{1/64}{1-1/16}=\frac1{60}.
\]
Writing \(b=2^{-1/2}\), retaining \(j=0,1,2\) in the denominator and
\(j=0,1,2,3\) in the numerator gives
\[
 \frac{\sum_{j\ge0}j2^{-j^2/2}}
      {\sum_{j\ge0}2^{-j^2/2}}
 <\frac{19b/16+1/2+1/60}{5/4+b}<\frac45,
\]
where the last inequality follows from \(b<71/100\). Finally,
\(2/3<\ln 2<5/7\) gives \(2+4/5<2/\ln 2\).

At \(\lambda=9{\ln 2}/2\), put \(t=2^{-1/8}\). The identity
\[
 \frac{9\ln 2}{2}i-\frac{\ln 2}{2}i^2
 =\frac{\ln 2}{8}\{81-(2i-9)^2\}
\]
gives
\[
 w_i(9{\ln 2}/2)=2^{81/8}t^{(2i-9)^2}.
\]
Thus the four weights, in the order
\(i=2,3,4,5\), are proportional to

\[
 t^{25},\quad t^9,\quad t,\quad t.
\]

The numerator of \(M_{S_4}(9{\ln 2}/2)-4\) is therefore

\[
 t-t^9-2t^{25}=t(1-t^8-2t^{24})=t/4>0,                 \tag{5.5}
\]

and the denominator is positive.  Hence this mean is greater than four,
whereas \(T_0\le1+2/{\ln 2}<4\).  Since \(M_{S_4}\) is strictly
increasing, the two mean comparisons give

\[
 2{\ln 2}<\lambda_4<9{\ln 2}/2.                                     \tag{5.6}
\]

For a tilt \(\lambda\), define the omitted low- and high-deficit ratios

\[
 L(\lambda):=
 \frac{\sum_{i=-1}^{1}w_i(\lambda)}{Z_{S_4}(\lambda)},
 \qquad
 H(\lambda):=
 \frac{\sum_{i\ge6}w_i(\lambda)}{Z_{S_4}(\lambda)}.
\]

Let \(M_A(\lambda)\) denote the mean of \(i\) under the weights
\(w_i(\lambda)\) restricted to an index set \(A\). Differentiation gives
\[
 \frac{d}{d\lambda}\ln L(\lambda)
 =M_{\{-1,0,1\}}(\lambda)-M_{S_4}(\lambda)<0,
\]
because the first mean is at most \(1\) and the second is at least \(2\).
Similarly,
\[
 \frac{d}{d\lambda}\ln H(\lambda)
 =M_{\{6,7,\ldots\}}(\lambda)-M_{S_4}(\lambda)>0,
\]
because the two means are at least \(6\) and at most \(5\), respectively.
Direct summation and Gaussian-tail bounds give

\[
 L(2{\ln 2})<\frac{51}{100},\quad H(3{\ln 2})<\frac1{50},\quad
 L(3{\ln 2})<\frac3{25},\quad H(9{\ln 2}/2)<\frac14.                 \tag{5.7}
\]

Here are explicit checks. At \(2{\ln 2}\), with \(b=2^{-1/2}\),

\[
 L(2{\ln 2})=\frac{b+1/4+b/16}{1+b+1/4+b/16}<51/100.
\]

At \(\lambda=3\ln 2\),
\[
 w_i(3\ln 2)
 =\exp\!\left((\ln 2)\left(3i-\frac{i^2}{2}\right)\right)
 =2^{9/2}2^{-(i-3)^2/2}.
\]
Set \(b=2^{-1/2}\). After canceling the common factor \(2^{9/2}\),
the weight at index \(i\) is \(b^{(i-3)^2}\). Thus the four
\(S_4\)-weights, for \(i=2,3,4,5\), are
\[
 b,\quad 1,\quad b,\quad b^4,
\]
and
\[
 2^{-9/2}Z_{S_4}(3\ln 2)=1+2b+b^4=\frac54+2b.
\]
For the omitted low indices \(i=-1,0,1\), the weights are
\[
 b^{16},\quad b^9,\quad b^4
 =\frac1{256},\quad\frac b{16},\quad\frac14.
\]
For the omitted high indices \(i\ge6\), they are
\[
 b^9+b^{16}+\sum_{q\ge5}b^{q^2}.
\]
The first term in the remaining sum is \(b^{25}=b/4096\), and the
ratio between successive terms is at most \(b^{11}<1/32\). Therefore
\[
 \sum_{q\ge5}b^{q^2}
 \le\frac{b^{25}}{1-1/32}<\frac1{3968}.
\]
It follows that
\[
 L(3\ln 2)
 =\frac{1/256+b/16+1/4}{5/4+2b}<\frac3{25},
\]
and
\[
 H(3\ln 2)
 \le\frac{b/16+1/256+1/3968}{5/4+2b}<\frac1{50},
\]
where the final inequalities follow from \(7/10<b<71/100\).

Finally, at \(9\ln 2/2\) the normalized exponent is
\((2i-9)^2\). Dividing all weights by the common \(S_4\) weight \(t\),
the high indices \(i=6,7,8\) contribute \(t^8,t^{24},t^{48}\).
From \(i=8\) onward the ratio between successive terms is at most
\(1/16\), so the tail beginning with \(t^{48}\) is less than \(1/60\).
The normalized \(S_4\) denominator is \(2+t^8+t^{24}\). Hence
\(H(9\ln 2/2)<1/4\) follows from

\[
 3(t^8+t^{24})+4/60<2,                                  \tag{5.8}
\]

because \(t^8=1/2\) and \(t^{24}=1/8\).

If \(\lambda_4\le 3{\ln 2}\), monotonicity and (5.7) give
\(L(\lambda_4)+H(\lambda_4)<53/100\); if
\(\lambda_4\ge 3{\ln 2}\), they give
\(L(\lambda_4)+H(\lambda_4)<37/100\).
By the dual representation above,

\[
 \mathcal F_{S_4}(T_0)
 =\ln Z_{S_4}(\lambda_4)-\lambda_4T_0,
\]

whereas evaluating the \(S_+\) dual function at the same parameter gives

\[
 \mathcal F_{S_+}(T_0)
 \le \ln Z_{S_+}(\lambda_4)-\lambda_4T_0.
\]

Subtracting these two relations yields

\[
 \begin{aligned}
 D_4(\delta)
 &\le \ln\frac{Z_{S_+}(\lambda_4)}{Z_{S_4}(\lambda_4)}\\
 &=\ln\!\bigl(1+L(\lambda_4)+H(\lambda_4)\bigr)
 <\ln(153/100).
 \end{aligned}
\]

Since \(S_4\) is a subset of \(S_+\), the loss is
nonnegative. Subtracting from \(\ln 2\) completes the proof.
\(\square\)

The following \(2^k\) gain is the one-partition identity isolated by
\citet[Proposition~6]{heckel-2025-difference}.  We repeat its short
argument because it is an input to the new root calculation.  For a fixed
partition into \(k\) classes, assigning each class the declaration
``independent'' or ``complete'' multiplies its first moment by \(2^k\):
every declaration prescribes all internal edge bits, and all \(2^k\)
declarations have probability \(2^{-\sum \binom{u}{2}}\).  The declarations
are disjoint here because all four allowed class sizes tend to infinity
(in particular, there are no singleton classes).
Let
\(r_4^{\mathrm{co}}=r_{S_4,{\ln 2}}\)
be the unique corridor zero of

\[
 L_{S_4}(n,k)+{\ln 2}k.                                       \tag{5.9}
\]

Put
\[
 \Phi_n(k):=L_{S_4}(n,k)+{\ln 2}k.
\]
Then \(\Phi_n(r_4^{\mathrm{co}})=0\) by definition, while at
\(k=r_+\), Lemma 3.1 and (5.2) give

\[
 L_{S_4}(n,r_+)+{\ln 2}r_+
 =r_+\{{\ln 2}-D_4(\delta)+o(1)\}.                             \tag{5.10}
\]

Both roots lie in the corridor (3.6), so the whole interval between them
is inside a fixed corridor.  The mean-value theorem therefore gives a
point \(\xi_n\in(r_4^{\mathrm{co}},r_+)\) such that
\[
 \Phi_n(r_+)-\Phi_n(r_4^{\mathrm{co}})
 =\Phi_n'(\xi_n)(r_+-r_4^{\mathrm{co}}).
\]
Now \(r_+=({\ln 2}/2+o(1))n/{\ln n}\), and (3.7), uniformly at
\(\xi_n\), gives
\[
 \Phi_n'(\xi_n)=\frac2{\ln 2}(\ln n)^2
                 +O((\ln n)(\ln\ln n)).
\]
Substituting this and (5.10) into the preceding identity yields

\[
 r_+-r_4^{\mathrm{co}}
 =\left(\frac{(\ln 2)^2}{4}\{{\ln 2}-D_4(\delta)\}+o(1)\right)
   \frac n{(\ln n)^3}.                                        \tag{5.11}
\]

In particular, for all sufficiently large \(n\),

\[
 r_+-r_4^{\mathrm{co}}\ge\frac{(\ln 2)^2\gamma_4}{8}\frac n{(\ln n)^3}.     \tag{5.12}
\]

Choose the midpoint integer

\[
 k_{\mathrm{co}}=\left\lceil\frac{r_4^{\mathrm{co}}+r_+}{2}\right\rceil. \tag{5.13}
\]

We call the resulting profile the \emph{exact midpoint profile}: its total
number of parts is the rounded midpoint between the signed four-size root
\(r_4^{\mathrm{co}}\) and the unrestricted chromatic root \(r_+\).  The
type counts are chosen next so that both finite-\(n\) conservation
constraints hold exactly.

At this exact \(k\), let \(p_i^{(n)}\) be the exact
finite-\(n\) maximizer of \(L_{S_4}\).  Let
\(\lambda_{n,\mathrm{def}}\) denote the Lagrange multiplier for the
deficit-mean constraint, with the sign convention that its contribution
is \(e^{\lambda_{n,\mathrm{def}}i}\).  The Lagrange equations and the
decomposition (3.11) then give

\[
 p_i^{(n)}\ \propto\ d_{\alpha-i}^{-1}e^{\lambda_{n,\mathrm{def}}i}
 \ \propto\ e^{\widehat\lambda_ni+h_n(i)},\qquad
 \widehat\lambda_n:=B_n+\lambda_{n,\mathrm{def}}.          \tag{5.14}
\]

Thus \(\lambda_{n,\mathrm{def}}\) is the multiplier before the affine
term \(B_ni\) is absorbed, while \(\widehat\lambda_n\) is the effective
tilt in the limiting Gaussian coordinates.  (If a size-coordinate
multiplier \(\tau_n\) is used instead, then
\(\lambda_{n,\mathrm{def}}=-\tau_n\).)
The target deficit mean is \(\alpha-n/k_{\mathrm{co}}\). Lemma 3.1 makes these proportions
converge uniformly to (5.1), with
\(\widehat\lambda_n\to\lambda_4(T_0)\). Equation (5.6) also shows
directly that the four limiting weights differ by at most a fixed
factor. Thus

\[
 \min_{2\le i\le5}p_i^{(n)}\ge c_p>0                    \tag{5.15}
\]

uniformly in the phase.

Round \(k_{\mathrm{co}}p_i^{(n)}\) to preliminary integers \(\widetilde k_i\).
Define the signed constraint errors

\[
 e_0=\sum_i\widetilde k_i-k_{\mathrm{co}},\qquad
 e_1=\sum_i i\widetilde k_i-(\alpha k_{\mathrm{co}}-n),
\]

so \(e_0,e_1=O(1)\), and add

\[
 \Delta k_2=e_1-3e_0,\qquad
 \Delta k_3=2e_0-e_1.                                   \tag{5.16}
\]

Indeed, \(\Delta k_2+\Delta k_3=-e_0\) and
\(2\Delta k_2+3\Delta k_3=-e_1\). This enforces both constraints
exactly, changes only \(O(1)\) counts, and preserves positivity by
(5.15).  Put
\(\mathbf k^*=(k_{\mathrm{co}}p_i^{(n)})_{i=2}^5\) and let
\(\Delta^{\mathrm{tot}}=\mathbf k-\mathbf k^*\) be the total rounding
and correction displacement.  It has bounded norm and satisfies both
homogeneous constraint equations, because \(\mathbf k^*\) and
\(\mathbf k\) obey the same two constraints.  Thus it is tangent to the
feasible affine plane.  At the exact optimizer the linear term vanishes, while
the Hessian of the entropy term is
\(-\operatorname{diag}(1/k_i)\).  Hence

\[
 L_{\mathbf k^*+\Delta^{\mathrm{tot}}}
 -L_{\mathbf k^*}
 =O\!\left(\sum_i\frac{(\Delta_i^{\mathrm{tot}})^2}{k_i^*}\right)
 =O(1/k_{\mathrm{co}}).
\]

The preceding rounding and correction have already produced the exact
integer vector

\[
 \mathbf k=(k_2,k_3,k_4,k_5),\qquad
 u_i=\alpha-i,\qquad k_i=\Theta(n/{\ln n}).                    \tag{5.17}
\]

Applying (1.2) to the factorials in the exact signed first moment of this
vector gives

\[
 \Exp{Z_{\mathbf k}^{\mathrm{sgn}}}
 =2^{k_{\mathrm{co}}}\frac{n!}{\prod_i(u_i!)^{k_i}k_i!}
   2^{-\sum_i k_i\binom{u_i}{2}},                        \tag{5.18}
\]

and

\[
 \ln \Exp{Z_{\mathbf k}^{\mathrm{sgn}}}
 =L_{S_4}(n,k_{\mathrm{co}})+{\ln 2}k_{\mathrm{co}}+O({\ln n})
 \ge c_Zk_{\mathrm{co}}                                          \tag{5.19}
\]

for a phase-independent \(c_Z>0\). Indeed, the midpoint is at distance
at least \((r_+-r_4^{\mathrm{co}})/2-O(1)\) to the right of
\(r_4^{\mathrm{co}}\).  Integrating the positive derivative (3.7) over
that interval and using (5.12) gives the scale calculation
\[
 \frac{n}{(\ln n)^3}\cdot(\ln n)^2
 \asymp\frac n{\ln n}\asymp k_{\mathrm{co}}.
\]
Consequently,
\[
 L_{S_4}(n,k_{\mathrm{co}})+{\ln 2}k_{\mathrm{co}}
 \ge c_Zk_{\mathrm{co}},
\]
after decreasing \(c_Z\) if necessary. Every
\(u_i\) tends to infinity, so the independent and complete
declarations of one part are disjoint. Thus
\(Z_{\mathbf k}^{\mathrm{sgn}}>0\)
implies an actual cocoloring with \(k_{\mathrm{co}}\) parts.

Finally, (4.1), (5.12), and (5.13) give

\[
 k_\chi^- -k_{\mathrm{co}}
 \ge\left(\frac{(\ln 2)^2\gamma_4}{16}-o(1)\right)
       \frac n{(\ln n)^3}.                                    \tag{5.20}
\]

\section{Exact signed second-moment
representation}\label{exact-signed-second-moment-representation}

For two partitions sharing exactly \(\ell\) whole classes,
\citet[Proposition~6]{heckel-2025-difference} bounded the number of joint
sign declarations by \(2^{2k-\ell}\), thereby transferring selected
ordinary-coloring overlap bounds to cocolorings.  Lemma 6.1 below does
not import that inequality.  It keeps the entire overlap matrix and counts
the compatible signs exactly: compatibility is controlled by the
components of the graph of cells of size at least two.  Whole common
classes appear as one special case of this component formula.

Temporarily label all slots within each of the four types. This
multiplies the unordered witness by the deterministic factor
\(\prod_i k_i!\) and hence does not change its normalized second
moment. Choose two independent uniform ordered partitions of the
profile. If their slots are indexed by \(a\) and \(b\), put

\[
 r_{ab}=|V_a\cap W_b|.                                   \tag{6.1}
\]

The overlap matrix has the exact law

\[
 p(r)=\frac{\prod_a s_a!\prod_b t_b!}
            {n!\prod_{a,b}r_{ab}!},                      \tag{6.2}
\]

where here both degree lists are the multiset given by (5.17).
Equivalently, place \(s_a\) stubs at each row slot and
\(t_b\) stubs at each column slot and take a uniform perfect
matching of the two \(n\)-stub sets.

Let \(H(r)\) be the simple bipartite graph whose edges are cells
with \(r_{ab}\ge 2\), omitting isolated slot vertices.
Write

\[
 W=\sum_{a,b}\binom{r_{ab}}2,\qquad
 \beta(H)=|E(H)|-|V(H)|+c(H).                             \tag{6.3}
\]

\phantomsection
\addcontentsline{toc}{subsection}{Lemma 6.1 (exact sign sum)}
\label{lemma-6.1-exact-sign-sum}
\begin{lemmabox}{Lemma 6.1 (exact sign sum)}

The normalized local factor of the pair of partitions is

\[
 A_\zeta(r)=2^{W+c(H)-|V(H)|}
 =\left(\prod_{a,b}g(r_{ab})\right)2^{\beta(H)},          \tag{6.4}
\]

where

\[
 g(0)=g(1)=g(2)=1,\qquad
 g(x)=2^{\binom x2-1}\quad(x\ge3).                       \tag{6.5}
\]

Consequently

\[
 \frac{\Exp{(Z_{\mathbf k}^{\mathrm{sgn}})^2}}
      {\Exp{Z_{\mathbf k}^{\mathrm{sgn}}}^2}
 =\sum_{r\in\mathcal R(\mathbf s,\mathbf t)}
      p(r)A_\zeta(r).                                    \tag{6.6}
\]

where \(\mathcal R(\mathbf s,\mathbf t)\) is the finite set of
nonnegative integer matrices with the prescribed row margins \(\mathbf s\)
and column margins \(\mathbf t\).

\end{lemmabox}

\paragraph{Proof.}\label{proof-3}

Let
\(B=\sum_a \binom{s_a}{2}\),
which is the number of internal edge bits prescribed by one partition. A
row sign and a column sign are compatible on a cell of size at least two
exactly when they agree. Therefore signs must be constant on every
component of \(H\). There are

\[
 2^{2k_{\mathrm{co}}-|V(H)|+c(H)}
\]

compatible pairs of sign assignments, including the free signs on slots
isolated from \(H\). For each compatible pair, \(W\) edge bits
are prescribed twice, so the probability of all constraints is
\(2^{-(2B-W)}\). Division by the square of the one-partition
signed probability
\((2^{k_{\mathrm{co}}}2^{-B})^2\)
gives

\[
 \frac{2^{2k_{\mathrm{co}}-|V(H)|+c(H)}2^{-(2B-W)}}
      {(2^{k_{\mathrm{co}}}2^{-B})^2}
 =2^{W+c(H)-|V(H)|},
\]

which is the first expression in (6.4).

Formula (6.4) separates two effects that will be estimated by different
methods. The product of the \(g(r_{ab})\) is purely local and is trivial for
cell sizes at most two; the factor \(2^{\beta(H)}\) is topological and counts
the independent cycle choices of the support graph.

For every support edge, split \(\binom{r}{2}\) as
\((\binom{r}{2}-1)+1\). Since
\(|E|-|V|+c=\beta\),

\[
 \prod_{a,b}g(r_{ab})=2^{W-|E(H)|},
\]

and multiplying by \(2^{\beta(H)}\) gives
the second expression. Averaging over (6.2) proves (6.6). \(\square\)

The identity

\[
 2^{\beta(H)}
 =\#\{F\subseteq E(H):\deg_F(v)\text{ is even for every }v\} \tag{6.7}
\]

will be used repeatedly. It is the elementary fact that the even
subgraphs form the binary cycle space of dimension \(\beta(H)\).

We also need one exact configuration-model estimate.

\phantomsection
\addcontentsline{toc}{subsection}{Lemma 6.2 (joint prescribed-cell bound)}
\label{lemma-6.2-joint-prescribed-cell-bound}
\begin{lemmabox}{Lemma 6.2 (joint prescribed-cell bound)}

In a bipartite configuration model of total degree \(m_0\), with
row degrees \(d_a\) and column degrees
\(d'_b\), prescribe distinct cells \(D\) and
demands \(x_{ab}\ge 1\). Put

\[
 x=\sum_Dx_{ab},\quad D_a=\sum_bx_{ab},\quad
 D'_b=\sum_ax_{ab}.
\]

In the feasible case

\[
 x\le m_0,\qquad D_a\le d_a\ \ (\forall a),\qquad
 D'_b\le d'_b\ \ (\forall b),
\]

we have

\[
 \Prob{r_{ab}\ge x_{ab}\ \text{for every }ab\in D}
 \le\frac{\prod_a(d_a)_{D_a}\prod_b(d'_b)_{D'_b}}
          {(m_0)_x\prod_{ab\in D}x_{ab}!}.               \tag{6.8}
\]

If any feasibility condition fails, the prescribed event is empty and is
handled separately. In particular, in all cases with \(m_0>0\), with
\(\mathrm e:=\exp(1)\) and
\(\theta_{ab}=\mathrm e\, d_a d'_b/m_0\),
its probability is at most

\[
 \prod_{ab\in D}\frac{\theta_{ab}^{x_{ab}}}{x_{ab}!}.  \tag{6.9}
\]

\end{lemmabox}

\paragraph{Proof.}\label{proof-4}

If a feasibility condition fails, no matching realizes all demands. In the
feasible case, for each row \(a\), assigning disjoint row stubs to its
demanded cells gives
\((d_a)_{D_a}/\prod_bx_{ab}!\) choices.  The analogous column choice
gives \((d'_b)_{D'_b}/\prod_ax_{ab}!\) choices, and there are
\(x_{ab}!\) bijections between the selected stubs inside cell \(ab\).
Thus the total number of demand witnesses is

\[
 \frac{\prod_a(d_a)_{D_a}\prod_b(d'_b)_{D'_b}}
      {\prod_{ab\in D}x_{ab}!}.
\]

Any fixed witness occurs with probability \(1/(m_0)_x\), because after
its \(x\) prescribed pairs are exposed the remaining matching is uniform
on the unused stubs.  A union bound proves (6.8).

It remains to simplify the global denominator.  For \(1\le x\le m_0\),

\[
 \binom{m_0}{x}
 =\prod_{j=0}^{x-1}\frac{m_0-j}{x-j}
 \ge\left(\frac{m_0}{x}\right)^x,
 \qquad
 x!\ge\left(\frac{x}{\mathrm e}\right)^x.
\]

The factorial inequality follows, for example, from
\(\sum_{j=1}^x\ln j\ge\int_1^x\ln t\,dt=x\ln x-x+1\).
Consequently

\[
 (m_0)_x=x!\binom{m_0}{x}\ge(m_0/\mathrm e)^x,          \tag{6.10}
\]

with the case \(x=0\) immediate.  Combining (6.10) with
\((d)_r\le d^r\) proves (6.9). Notice that (6.8) retains the single global falling
factorial before (6.9) is taken. \(\square\)

\section{Exact partial diagonals}\label{exact-partial-diagonals}

\paragraph{Purpose of this section.}
The exact overlap representation of Section~6 still permits two partitions
to share whole classes. The definitions below assign a nonnegative marked
weight to each such common subprofile. We prove that the total marked
contribution is \(1+o(1)\), uniformly through the phase. This
partial-diagonal estimate is the input used in Section~8 to control the
dense transportation sum. The counting identities in (7.1)--(7.4) are
finite-\(n\) identities; asymptotic estimates are explicitly labeled below.

A \emph{partial diagonal} is a specified collection of whole classes that
the two partitions have in common.  It is marked: one overlap may contain
several such collections and then contributes to every corresponding
nonnegative marked term.  This intentional overcount is what makes the
subsequent summation an upper-bound device.

For a selected common subprofile \(\ell=(\ell_i)\), put

\[
 \ell_\bullet=\sum_i\ell_i,\qquad
 m=\sum_i u_i\ell_i,\qquad
 b_\ell=\sum_i\binom{u_i}{2}\ell_i.                           \tag{7.1}
\]

Here \(0\le\ell_i\le k_i\) coordinatewise.  Since the full profile has
\(\sum_i u_i k_i=n\), this automatically gives \(m\le n\); thus every
factorial in the counting formula below has its ordinary combinatorial
meaning.

\paragraph{Exact marked decomposition.}
The signed first moment of such partial partitions occupying any
\(m\) vertices of \([n]\) is

\[
 Y_{\ell}^{\mathrm{sgn}}
 =2^\ell_\bullet\frac{n!}{(n-m)!\prod_i(u_i!)^{\ell_i}\ell_i!}
       2^{-b_\ell}.                                           \tag{7.2}
\]

For each type \(i\), choose the \(\ell_i\) marked slots in the first
partition and the \(\ell_i\) marked slots in the second, giving
\(\binom{k_i}{\ell_i}^2\) choices. The common blocks themselves have the
signed first-moment weight \(Y_\ell^{\mathrm{sgn}}\). In the normalized
second moment they are counted once rather than independently in each
marginal, so their contribution is the reciprocal of this weight.
Therefore the exact \emph{marked} common-subprofile weight is

\[
 D(\ell)=\frac{\prod_i\binom{k_i}{\ell_i}^2}
                 {Y_\ell^{\mathrm{sgn}}}.                         \tag{7.3}
\]

Different marked choices need not be disjoint: an overlap containing several
common blocks contributes to each corresponding marked choice. The sums below
are sums of these nonnegative marked weights, so no disjointness is used.
In particular,
\(D(0)=1\) and \(D(k)=1/\Exp{Z_k^{\mathrm{sgn}}}\).
To add one common block of type \(i\), first note that

\[
 \frac{\binom{k_i}{\ell_i+1}^2}{\binom{k_i}{\ell_i}^2}
 =\frac{(k_i-\ell_i)^2}{(\ell_i+1)^2},
\]

while (7.2) gives

\[
 \frac{Y_{\ell+e_i}^{\mathrm{sgn}}}
      {Y_\ell^{\mathrm{sgn}}}
 =\frac{2(n-m)_{u_i}}
        {u_i!(\ell_i+1)2^{\binom{u_i}{2}}}
 =\frac{2\mu_{u_i}(n-m)}{\ell_i+1}.
\]

Dividing these two ratios gives

\[
 \frac{D(\ell+e_i)}{D(\ell)}
 =\frac{(k_i-\ell_i)^2}
        {2(\ell_i+1)\mu_{u_i}(n-m)}.                     \tag{7.4}
\]

At the opposite endpoint, put \(h=k-\ell\),
\(v=\sum_i u_i h_i\), and let \(Z_h^{\mathrm{sgn}}(v)\) be the
signed first moment of a complete profile \(h\) on \(v\)
vertices. Exact factorial cancellation gives

\[
 D(k-h)=\frac{B(h)}{\Exp{Z_k^{\mathrm{sgn}}}},\qquad
 B(h)=\left(\prod_i\binom{k_i}{h_i}\right)Z_h^{\mathrm{sgn}}(v), \tag{7.5}
\]

and

\[
 \frac{\binom{k_i}{h_i+1}}{\binom{k_i}{h_i}}
 =\frac{k_i-h_i}{h_i+1},\qquad
 \frac{Z_{h+e_i}^{\mathrm{sgn}}(v+u_i)}
      {Z_h^{\mathrm{sgn}}(v)}
 =\frac{2\mu_{u_i}(v+u_i)}{h_i+1}.
\]

Therefore

\[
\frac{B(h+e_i)}{B(h)}
 =\frac{2(k_i-h_i)\mu_{u_i}(v+u_i)}{(h_i+1)^2}.          \tag{7.6}
\]

Equations (7.2)--(7.6) are exact finite-\(n\) counting identities.  We now
use them only through uniform asymptotic bounds on the three ranges of
selected mass.

In Heckel's admissible-phase argument, the analogous control of partial
profiles enters through the tame-coloring estimates of
\citet[Lemmas~5.1--5.3, 6.3--6.5, and~7.20]{heckel-panagiotou-2023},
as used in \citet[Proposition~5]{heckel-2025-difference}; their hypotheses are the
source of the lower \(\mu_\alpha\)-window in that theorem.  The next
lemma is the replacement needed here.  It is proved directly for the
four-size signed profile, including both corners and every intermediate
mass, and no tame-profile theorem is invoked.

\phantomsection
\addcontentsline{toc}{subsection}{Lemma 7.1 (all common subprofiles)}
\label{lemma-7.1-all-common-subprofiles}
\begin{lemmabox}{Lemma 7.1 (all common subprofiles)}

For the exact midpoint profile,

\[
 \sum_{0\le\ell_i\le k_i}D(\ell)=1+o(1).               \tag{7.7}
\]

\end{lemmabox}

The proof has three ranges, classified by the vertex mass occupied by the
marked common classes. The recurrence (7.4) controls the empty corner, the
continuous rate function controls the central range, and the reverse
recurrence (7.6) controls the full corner. The ranges overlap only at their
boundaries and together exhaust every common subprofile.

\paragraph{Proof: the empty corner.}\label{proof-the-empty-corner}

Let

 \[
  \eta=\frac{{\ln\ln n}}{32{\ln n}},\qquad
 \xi_i=\frac{k_i^2}{2\mu_{u_i}(n)},\qquad
 \Xi_{\mathrm{empty}}=\sum_i\xi_i.                               \tag{7.8}
 \]

Equations (2.4), (2.8), and \(k_i=\Theta(n/{\ln n})\) imply

\[
 \Xi_{\mathrm{empty}}=O((\ln n)^{-(2/{\ln 2}-1/2)})=o(1).        \tag{7.9}
\]

If the final selected mass is at most \(\eta n\), every
intermediate mass \(m'\) in (7.4) satisfies

\[
 \frac{\mu_{u_i}(n)}{\mu_{u_i}(n-m')}
 \le(1-2\eta)^{-u_i}\le (\ln n)^{3/8}.                        \tag{7.10}
\]

Iterating (7.4), in any order, yields

\[
 D(\ell)\le\prod_i\frac{((\ln n)^{3/8}\xi_i)^{\ell_i}}
                               {\ell_i!}.
\]

Thus

\[
 1\le\sum_{m\le\eta n}D(\ell)
 \le\exp((\ln n)^{3/8}\Xi_{\mathrm{empty}})=1+o(1).       \tag{7.11}
\]

\paragraph{Proof: the central range.}\label{proof-the-central-range}

Write

\[\begin{gathered}
p_i=k_i/k_{\mathrm{co}},\quad y_i=\ell_i/k_{\mathrm{co}},\quad
 Y=\sum_i y_i,\quad I=\sum_i i y_i,\quad\\
z_i=p_i-y_i,\quad R=\sum_i z_i=1-Y,\quad
 I_r=\sum_i i z_i=T-I.
\end{gathered}
\tag{7.12}\]

The residual vertex fraction is

\[
 \rho=\frac{n-m}{n}
 =R+\frac{I-TY}{\alpha-T}.                              \tag{7.13}
\]

\paragraph{Uniform asymptotic estimate: Stirling control.}
Before applying Stirling, (7.2) and (7.3) give the exact logarithmic
identity
\[
\begin{split}
\ln D(\ell)={}&
2\sum_i\{\ln(k_i!)-\ln(\ell_i!)-\ln((k_i-\ell_i)!)\}\\
&-\ell_\bullet\ln 2-\ln(n!)+\ln((n-m)!)
 +\sum_i\ell_i\ln(u_i!)+\sum_i\ln(\ell_i!)
 +(\ln 2)b_\ell.
\end{split}
\]
Applying Stirling to this identity, with the convention
\(0\ln0=0\), and using \(n=k_{\mathrm{co}}(\alpha-T)\), gives uniformly

\[
\begin{split}
 \ln D(\ell)={}&n\rho\ln\rho\\
 &+k_{\mathrm{co}}\sum_i\{2p_i\ln p_i-2(p_i-y_i)\ln(p_i-y_i)
                  -y_i\ln y_i-y_i+y_iE_i\}+O({\ln n}),        \end{split}
\tag{7.14}
\]

where

\[
 E_i=\ln k_{\mathrm{co}}+\ln(u_i!)+u_i-u_i{\ln n}
       +{\ln 2}\binom{u_i}{2}-{\ln 2}.                               \tag{7.15}
\]

Since \(u_{i+1}=u_i-1\), exact subtraction in (7.15) gives

\[
 \begin{aligned}
 E_{i+1}-E_i
 &=\ln n-\ln(\alpha-i)-(\ln 2)(\alpha-i)+\ln 2-1\\
 &=-\frac{\ln 2}{2}\alpha+O(1).
 \end{aligned}                                                 \tag{7.16}
\]

For the second line, substitute
\((\ln 2)\alpha/2=\ln n-\ln\ln n+O(1)\) from (2.1) and
\(\ln(\alpha-i)=\ln\ln n+O(1)\), uniformly for \(2\le i\le5\).

Let \(\bar E=\sum p_i E_i\). Applying Stirling to the complete
moment (5.18) gives

\[
 -\frac1{k_{\mathrm{co}}}\ln \Exp{Z_k^{\mathrm{sgn}}}
 =\sum_i p_i\ln p_i-1+\bar E+o(1).                      \tag{7.17}
\]

The positive margin (5.19) bounds \(\bar E\) above.  Indeed, (7.17)
and \(\ln\Exp{Z_k^{\mathrm{sgn}}}\ge c_Zk_{\mathrm{co}}\)
give \(\bar E=O(1)\) from above, uniformly in the phase.  Since the
support consists of four consecutive indices, summing (7.16) over at
most three steps and using \(\sum_i p_i(i-T)=0\) gives

\[
 E_i=\bar E-\frac{{\ln 2}\alpha}{2}(i-T)+O(1)
 \qquad(2\le i\le5).
\]

Multiplying by \(y_i\) and summing yields

\[
 \sum_i y_iE_i\le\frac{{\ln 2}\alpha}{2}(TY-I)+O(Y).          \tag{7.18}
\]

Indeed,
\[
 \sum_i y_iE_i
 =Y\bar E-\frac{(\ln 2)\alpha}{2}(I-TY)+O(Y).
\]
Only an upper bound is needed in (7.18), and \(\bar E\le C\) uniformly;
therefore \(Y\bar E\le CY\), which explains its absorption into
\(O(Y)\).
The remaining entropy expression in (7.14) is \(O(Y \ln(\mathrm e/Y))\),
uniformly because every \(p_i\) is bounded below. If
\(\rho\ge 1/32\), then \(R\ge 1/64\) for
large \(n\). Equation (7.13) and \(2\le i,T\le5\) give
\[
 |\rho-R|
 =\frac{|I-TY|}{\alpha-T}=O(Y/\alpha).
\]
The derivative of \(x\ln x\) is bounded on the interval between
\(\rho\) and \(R\). The mean-value theorem, together with
\(n=k_{\mathrm{co}}(\alpha-T)\) and
\(|R\ln R|=O(1-R)=O(Y)\), therefore gives

\[
 n\rho\ln\rho=k_{\mathrm{co}}\alpha R\ln R+O(k_{\mathrm{co}}Y).           \tag{7.19}
\]

Consequently

\[
 \ln D(\ell)
 \le k_{\mathrm{co}}\alpha\Phi_T(z)
     +Ck_{\mathrm{co}}Y\ln(\mathrm e/Y)+C{\ln n},                       \tag{7.20}
\]

where

\[
 \Phi_T(z)=R\ln R+\frac {\ln 2}2(I_r-TR).                     \tag{7.21}
\]

\paragraph{Uniform asymptotic estimate: rate negativity.}
This rate is uniformly negative away from its two corners. Indeed,

\[
 I_r-TR\le(5-T)R,\qquad
 I_r-TR=\sum_i(T-i)y_i\le(T-2)(1-R),                    \tag{7.22}
\]

and hence

\[
 \Phi_T\le R\{\ln R+{\ln 2}(5-T)/2\},
 \quad
 \Phi_T\le R\ln R+{\ln 2}(T-2)(1-R)/2.                       \tag{7.23}
\]

Here is the numerical check used in this split.  On
\(1/64\le R\le47/100\), add \(Y/5000=(1-R)/5000\) to the first
bound in (7.23).  The resulting function

\[
 R\ln R+\frac{{\ln 2}}2(5-T)R+\frac{1-R}{5000}
\]

is convex in \(R\).  Its largest possible coefficient occurs at
\(T=2/{\ln 2}\). Using
\[
 0.6931<\ln 2<0.6932,\qquad
 \ln 64>4.1588,\qquad \ln(100/47)>0.7550,
\]
its values at the two endpoints are bounded above by
\[
 -0.0533\quad\text{at }R=1/64,\qquad
 -0.0102\quad\text{at }R=47/100.
\]
Convexity therefore gives \(\Phi_T\le-Y/5000\) throughout this interval.

On \(47/100\le R\le1\), use the second bound in (7.23) and
\(T\le1+2/{\ln 2}\).  After adding \(Y/200\), it is at most

\[
 R\ln R+(1-{\ln 2}/2+1/200)(1-R)
\]

This function is convex; the same numerical bounds give a value below
\(-0.0058\) at \(R=47/100\), while its value at \(R=1\) is \(0\).
It follows that \(\Phi_T\le-Y/200\) on the second interval.
Thus, whenever

\[
 m>\eta n,\qquad n-m>n/32,                               \tag{7.24}
\]

we have \(Y\ge\eta/2\) for all sufficiently large \(n\). Indeed,
\[
 \frac mn=\frac{\alpha Y-I}{\alpha-T}
 \le\frac{\alpha Y}{\alpha-T},
\]
so \(m>\eta n\) implies
\[
 Y>\eta\frac{\alpha-T}{\alpha}\ge\frac{\eta}{2}.
\]
The leading term is at most
\(-k_{\mathrm{co}}\alpha Y/5000\), whereas
\(\ln(\mathrm e/Y)=O({\ln\ln n})\), \({\ln\ln n}/\alpha=o(1)\),
and \({\ln n}=o(k_{\mathrm{co}}\alpha\eta)\). It therefore dominates both error terms
in (7.20). Uniformly in this range,

\[
 D(\ell)\le e^{-c k_{\mathrm{co}}{\ln\ln n}}.                              \tag{7.25}
\]

There are only \((k_{\mathrm{co}}+1)^4\) vectors, so their total is
\(o(1)\).

\paragraph{Proof: the full corner.}\label{proof-the-full-corner}

The \emph{full corner} is the regime in which the selected common blocks
occupy all but a small residual vertex mass, namely \(v=n-m\le n/32\).
Uniformly for the
four sizes,
\(\mu_{u_i}(n)\le n^{6+o(1)}\)
by (2.2) and (2.8), while

\[
 \mu_{u_i}(v)\le\mu_{u_i}(n)(v/n)^{u_i}\le n^{-4+o(1)}. \tag{7.26}
\]

Here \(u_i=(2/{\ln 2}+o(1)){\ln n}\) and
\(32^{-u_i}=n^{-10+o(1)}\).
Consequently \(\mu_{u_i}(v+u_i)\le n^{-4+o(1)}\) at every intermediate
step while \(v\le n/32\).  Since \(k_i\le n\) and
\((h_i+1)^2\ge1\), the right side of (7.6) is at most
\(2n^{-3+o(1)}<1\).  Thus, in any order of adding residual blocks,
\(B(h)\) never increases from \(B(0)=1\), and hence \(B(h)\le1\).
Equations (5.19) and (7.5) give

\[
 D(k-h)\le e^{-c_Zk_{\mathrm{co}}}.                               \tag{7.27}
\]

Summing \((k_{\mathrm{co}}+1)^4\) residual vectors proves an \(o(1)\)
contribution. Together with (7.11) and (7.25), this proves (7.7).
\(\square\)

\section{Canonical high cells and dense endpoint
transportation}\label{canonical-high-cells-and-dense-endpoint-transportation}

\displayheading{Contract for this section}
The input is the exact overlap sum (6.6), after Section~7 has removed the
contribution of common whole classes.  The output is a bound for the sum of
the weights of all canonical high-cell skeletons, including unequal-type and
nonendpoint cells.  The local and cycle factors of the unexposed residual
configuration are deliberately deferred to Section~9.  Keeping those two
tasks separate is what prevents the same overlap from being charged both as a
high cell and as a residual attachment.

Put

\[
 u_{\max}=\alpha-2,\qquad R_0=\lfloor u_{\max}/2\rfloor.                \tag{8.1}
\]

In every overlap matrix, the cells of multiplicity greater than
\(R_0\) will be called \emph{high cells}. They form a matching: two high
cells in one row or column would use more than \(u_{\max}\) stubs. The
resulting skeleton is canonical: for a given overlap matrix it is exactly
the collection of all high cells, together with their multiplicities, and
there is no auxiliary choice. Write the
resulting canonical high-cell skeleton and its multiplicities as

\[
 \mathcal M=\{a_rb_r:1\le r\le h\},\qquad
 r_{a_rb_r}=j_r>R_0,\qquad J=\sum_rj_r.                  \tag{8.2}
\]

Selecting the exact stub pairs in these cells has incidence

\[
\pi(\mathcal M,j)=
 \frac{\prod_r(s_{a_r})_{j_r}(t_{b_r})_{j_r}}
      {(n)_J\prod_rj_r!}.                                \tag{8.3}
\]

The corresponding \emph{high-skeleton weight} is
\[
 w_{\mathrm{hi}}(\mathcal M,j)
 :=\pi(\mathcal M,j)\prod_{e\in\mathcal M}g(j_e).         \tag{8.3a}
\]
It contains the incidence probability and the signed local rewards of the
exposed high cells, but no residual local factor and no cycle-space factor.

Conditional on the exposed pairs, the remaining matching is a uniform
configuration model with total degree \(n-J\) and the residual row
and column degrees. It is restricted to put no further pair in a cell of
\(\mathcal M\) and no cell above \(R_0\). To see the cancellation
explicitly, let
\[
 s'_a=s_a-\sum_{r:a_r=a}j_r,\qquad
 t'_b=t_b-\sum_{r:b_r=b}j_r,
\]
and, for a final overlap matrix \(r\), let \(r'\) be obtained by replacing
each exposed entry \(r_{a_rb_r}=j_r\) by \(0\). The conditional probability
of \(r'\) is
\[
 \frac{\prod_a s'_a!\prod_b t'_b!}
      {(n-J)!\prod_{a,b}r'_{ab}!}.
\]
Multiplying this expression by (8.3), and using
\((s_a)_{s_a-s'_a}=s_a!/s'_a!\),
\((t_b)_{t_b-t'_b}=t_b!/t'_b!\), and
\((n)_J=n!/(n-J)!\), gives
\[
 \frac{\prod_a s_a!\prod_b t_b!}
      {n!\prod_{a,b}r_{ab}!},
\]
which is exactly (6.2). Thus every overlap matrix occurs exactly once.
The canonical choice of all cells above \(R_0\) is what prevents an overlap
matrix from being charged to two different skeletons.

We first sum the bare high-skeleton contribution. Index the four sizes
as

\[
 u_i=a-i,\qquad 0\le i\le3,\qquad a=\alpha-2.             \tag{8.4}
\]

\displayheading{Exact endpoint bookkeeping}
For a typed full-containment matrix \(\mathbf L=(\ell_{ij})\), put

\[
 r_i=\sum_j\ell_{ij},\quad c_j=\sum_i\ell_{ij},\quad
 x_{ij}=\min(u_i,u_j),\quad J(\mathbf L)=\sum_{ij}x_{ij}\ell_{ij}. \tag{8.5}
\]

Its exact incidence, including its local signed rewards but not residual
attachments, is

\[
 W(\mathbf L)=
 \frac{\prod_i(k_i)_{r_i}\prod_j(k_j)_{c_j}}
      {\prod_{ij}\ell_{ij}!}\frac1{(n)_{J(\mathbf L)}}
 \prod_{ij}\left[
  \frac{(u_i)_{x_{ij}}(u_j)_{x_{ij}}}{x_{ij}!}g(x_{ij})
 \right]^{\ell_{ij}}.                                   \tag{8.6}
\]

For a margin vector \(r\), define

\[
 D(r)=
 \frac{\prod_i(k_i)_{r_i}^2}{\prod_i r_i!}
 \frac{\prod_i[u_i!g(u_i)]^{r_i}}{(n)_{m_r}},\qquad
 m_r=\sum_i u_ir_i.                                     \tag{8.7}
\]

This is exactly the common-subprofile weight (7.3).

\phantomsection
\addcontentsline{toc}{subsection}{Lemma 8.1 (geometric-mean transportation comparison)}
\label{lemma-8.1-geometric-mean-transportation-comparison}
\begin{lemmabox}{Lemma 8.1 (geometric-mean transportation comparison)}

For every feasible \(\mathbf L\),

\[
 W(\mathbf L)\le\sqrt{D(r)D(c)}
 \frac{\sqrt{\prod_i r_i!c_i!}}{\prod_{ij}\ell_{ij}!}
 \prod_{ij}Q_{ij}^{\ell_{ij}},                          \tag{8.8}
\]

where \(Q_{ii}=1\) and, if the two sizes are \(t=s+d\),
\(1\le d\le 3\),

\[
 Q_{ij}=(n+1)^{d/2}\frac{\sqrt{(t)_d}}{d!}
       2^{-\{ds+\binom d2\}/2}
 \le\frac{\eta_n^d}{d!},                               \tag{8.9}
\]

with

\[
 \eta_n=2^{3/2}\sqrt{\frac{(n+1)a}{2^a}}
       =O((\ln n)^{3/2}/\sqrt n)=o(1).                         \tag{8.10}
\]

\end{lemmabox}

\begin{proof}\label{proof-5}

The estimate has two sources.  The first is a global falling-factorial
comparison: replacing a rectangular overlap table by its two one-sided
margin profiles costs only the small off-diagonal factor in (8.12).  The
second is a local comparison of unequal class sizes: a cell joining sizes
that differ by \(d\) carries the factor \(Q_{ij}\), and \(Q_{ij}\) is small
because the random graph penalty \(2^{-ds/2}\) dominates the polynomial
choices of the extra \(d\) vertices.

For integer \(x\), let \(f(x)=\ln(n)_x\) and interpolate
linearly between successive integers. Its successive slopes are
\(\ln(n-x)\), which decrease; therefore \(f\) is concave, and
every slope is at most \(\ln(n+1)\). Moreover,

\[
 \frac{m_r+m_c}{2}
 =J(\mathbf L)+\frac12\sum_{ij}|i-j|\ell_{ij}.                  \tag{8.11}
\]

Concavity first gives

\[
 \frac{f(m_r)+f(m_c)}2
 \le f\!\left(\frac{m_r+m_c}{2}\right).
\]

Using (8.11) and then moving from \(J(\mathbf L)\) along a distance
\(\frac12\sum_{ij}|i-j|\ell_{ij}\), with slope at most
\(\ln(n+1)\), gives

\[
 \frac{f(m_r)+f(m_c)}2-f(J(\mathbf L))
 \le\frac{\ln(n+1)}2\sum_{ij}|i-j|\ell_{ij}.
\]

Exponentiating this inequality gives

\[
 \frac{\sqrt{(n)_{m_r}(n)_{m_c}}}{(n)_{J(\mathbf L)}}
 \le(n+1)^{\frac12\sum_{ij}|i-j|\ell_{ij}}.            \tag{8.12}
\]

For \(t=s+d\), put \(b_u=u!g(u)\). A direct calculation gives

\[
 \frac{(t)_s(s)_s}{s!}g(s)=b_s\binom td,\qquad
 \frac{b_t}{b_s}=(t)_d2^{ds+\binom d2}.                 \tag{8.13}
\]

Thus the local factor divided by \(\sqrt{b_s b_t}\) is

\[
 \frac{\sqrt{(t)_d}}{d!}2^{-\{ds+\binom d2\}/2}.        \tag{8.14}
\]

Dividing (8.6) by \(\sqrt{D(r)D(c)}\), then using (8.12) and (8.14),
proves (8.8)--(8.9). Finally \(t\le a\),
\(s\ge a-3\), and

\[
 2^a=\Theta(n^2/(\ln n)^2)                                    \tag{8.15}
\]

uniformly in the phase. This proves (8.10).
\end{proof}

\phantomsection
\addcontentsline{toc}{subsection}{Lemma 8.2 (sum of all endpoint tables)}
\label{lemma-8.2-sum-of-all-endpoint-tables}
\begin{lemmabox}{Lemma 8.2 (sum of all endpoint tables)}

Let \(\mathcal P\) be the finite set of feasible one-sided margin vectors.
The first sum below is over all typed full-containment tables \(\mathbf L\) whose
row and column margins lie in \(\mathcal P\).  Then

\[
 \begin{aligned}
  \sum_{\mathbf L}W(\mathbf L)
 &\le \exp\{O(\eta_n k_{\mathrm{co}})\}
      \left(\sum_{\mathbf r\in\mathcal P}\sqrt{D(\mathbf r)}\right)^2\\
 &\le \exp\{O(\sqrt{n{\ln n}})+O({\ln n})\}
      \sum_{\mathbf r\in\mathcal P}D(\mathbf r)\\
 &=\exp\{O(\sqrt{n{\ln n}})+O({\ln n})\}.
 \end{aligned}
 \tag{8.16}
\]

\end{lemmabox}

\begin{proof}\label{proof-6}

This is the point at which the geometric-mean comparison becomes a genuine
sum over all endpoint tables.  For fixed margins we forget one set of margin
constraints at a time.  That enlargement is harmless because all summands are
nonnegative, and it converts the constrained table sum into a multinomial
sum.  Cauchy's inequality then restores symmetry between the two margins.

For fixed margins, write

\[
 A_{\mathbf L}=\frac{\prod_i r_i!}{\prod_{ij}\ell_{ij}!},\qquad
 C_{\mathbf L}=\frac{\prod_j c_j!}{\prod_{ij}\ell_{ij}!}.        \tag{8.17}
\]

Write \(Q^{\mathbf L}:=\prod_{ij}Q_{ij}^{\ell_{ij}}\).
Cauchy's inequality gives

\[
 \sum_{\mathbf L}\sqrt{A_{\mathbf L}C_{\mathbf L}}\,Q^{\mathbf L}
 \le\left(\sum_{\mathbf L}A_{\mathbf L}Q^{\mathbf L}\right)^{1/2}
      \left(\sum_{\mathbf L}C_{\mathbf L}Q^{\mathbf L}\right)^{1/2}.
                                                               \tag{8.18}
\]

All summands in (8.18) are nonnegative. In the first sum, remove the
column-margin constraints while retaining the fixed row margins; this
enlarges the summation domain and therefore gives an upper bound. The
unrestricted allocation of the \(r_i\) row objects among the column types is
then counted exactly by the multinomial theorem. The second inequality is
obtained symmetrically by removing the row-margin constraints while
retaining the column margins. Thus

\[
 \sum_{\mathbf L}A_{\mathbf L}Q^{\mathbf L}
 \le\prod_i\left(\sum_jQ_{ij}\right)^{r_i},
 \quad
 \sum_{\mathbf L}C_{\mathbf L}Q^{\mathbf L}
 \le\prod_j\left(\sum_iQ_{ij}\right)^{c_j}.
                                                               \tag{8.19}
\]

Every row and column sum of \(Q\) is at most \(1+C \eta_n\),
and the total margin is at most \(k_{\mathrm{co}}\). Hence the fixed-margin
cost is
\(\exp(O(\eta_n k_{\mathrm{co}}))\).
There are only \(O(k_{\mathrm{co}}^4)\) margin vectors, and

\[
  \left(\sum_{\mathbf r\in\mathcal P}\sqrt{D(\mathbf r)}\right)^2
  \le O(k_{\mathrm{co}}^4)\sum_{\mathbf r\in\mathcal P}D(\mathbf r).
                                                               \tag{8.20}
\]

Use Lemma 7.1 and
\(\eta_n k_{\mathrm{co}}=O(\sqrt{n{\ln n}})\)
to finish.
\end{proof}

\phantomsection
\addcontentsline{toc}{subsection}{Lemma 8.3 (all nonendpoint high multiplicities)}
\label{lemma-8.3-all-nonendpoint-high-multiplicities}
\begin{lemmabox}{Lemma 8.3 (all nonendpoint high multiplicities)}

After all feasible high-cell multiplicities
\(R_0<j\le\min(u_i,u_j)\) are included,
\[
 \sum_{(\mathcal M,j)\ \mathrm{feasible}}
 w_{\mathrm{hi}}(\mathcal M,j)
 \le \exp\!\left\{o\!\left(\frac{n}{(\ln n)^4}\right)\right\}.
\]

\end{lemmabox}

\begin{proof}\label{proof-7}

Replacing a high cell by full containment creates two qualitatively different
errors. A small deficit from the endpoint factorizes over the matching and is
summable cell by cell; a middle multiplicity must instead be estimated inside
the residual configuration model. The proof treats these mechanisms
separately and then recombines them through the exact canonical decomposition.

It may help to view the argument as a charging scheme.  Every high cell is
charged first to the nearest endpoint table.  If the cell is close to full
containment, the charge is paid immediately by the rapidly decreasing
one-cell series (8.25).  If it is not close to full containment, it is left in
the residual model and paid for by the joint threshold estimate (8.26a), or by
the crude small-residual bound (8.29a).  The no-double-counting point is that
the joint probability estimate is applied before the residual restrictions
are relaxed.  The proof is organized into five explicit steps: the canonical
near--middle split, the summation of near cells, the two residual-mass
branches for middle cells, and the final assembly.

\displayheading{Step I: the canonical near--middle split}

We first record why the following two ranges exhaust the canonical
high-skeleton sum. If a high cell joins slots of sizes \(m\) and \(m+d\),
where \(0\le d\le3\), write its multiplicity uniquely as \(j=m-e\).
Call the cell \emph{near} when \(e<m/4\), and \emph{middle} otherwise.
Because the high cells form a matching, replacing every near cell by the
endpoint multiplicity \(m\) produces a feasible typed full-containment
table \(\mathbf L\); the original near cells are recovered from
\(\mathbf L\) and their uniquely determined deficits \(e\). The remaining
high cells satisfy
\[
 R_0<j=m-e\le 3m/4\le3a/4.
\]
Thus every canonical high skeleton has a unique near part recorded by an
endpoint table and a residual middle part. In the upper bounds below we may
admit residual choices that do not themselves form a matching or satisfy
the no-near-cell restriction, but only after the relevant joint probability
has been bounded. Since all summands are nonnegative, these relaxations can
only increase the sum.

\displayheading{Step II: summing near-containment cells}
Let the smaller slot size be \(m\), the larger \(m+d\), with
\(0\le d\le 3\). Replacing the endpoint
multiplicity \(m\) by \(m-e\) has exact local ratio

\[
 R_{m,d}(e)=
 \frac{\binom me}{(d+1)\cdots(d+e)}
 2^{-em+e(e+1)/2}.                                      \tag{8.21}
\]

For many decorated cells, with total deficit \(Q\), the ratio of
the one global denominators is

\[
 \frac{(n)_{J_0}}{(n)_{J_0-Q}}
 =(n-J_0+Q)_Q\le n^Q.                                   \tag{8.22}
\]

Put
\[
 A_e:=n^eR_{m,d}(e),\qquad 0\le e<m/4.
\]
For \(0\le e<m/4\), its consecutive ratio is

\[
 \frac{n(m-e)2^{-m+e+1}}{(e+1)(d+e+1)}.                 \tag{8.23}
\]

The ratio of (8.23) at \(e+1\) to its value at \(e\) is

\[
 2\frac{m-e-1}{m-e}\frac{e+1}{e+2}
  \frac{d+e+1}{d+e+2},                                  \tag{8.24}
\]

which increases for \(e\le m/4\) and \(d\le3\). Here is a direct check
of the one-turn assertion. If \(r(e)\) denotes (8.23), interpolated for
real \(e\), then

\[
 \frac{d^2}{de^2}\ln r(e)
 =-\frac1{(m-e)^2}+\frac1{(e+1)^2}
   +\frac1{(d+e+1)^2}>0                                  \tag{8.24a}
\]

for \(0\le e\le m/4\) and all sufficiently large \(m\). Thus the
consecutive ratios are log-convex and their maximum is at an endpoint.
At \(e=0\) the ratio is \(O((\ln n)^3/n)\), by (8.15), and at
\(e=\lfloor m/4\rfloor\) it is \(n^{-1/2+o(1)}\). The series therefore
decreases geometrically and

\[
 \sum_{1\le e<m/4}n^eR_{m,d}(e)=O((\ln n)^3/n).                \tag{8.25}
\]

\displayheading{From one near cell to a full skeleton}
The passage from this per-cell estimate to the full skeleton sum is as
follows. For a full-containment table \(\mathbf L\), temporarily distinguish its
endpoint cells, write \(\operatorname{cells}(\mathbf L)\) for the resulting
multiset, and let \(\mathcal S(\mathbf L)\) be all near-containment
skeletons obtained by assigning to each such cell either \(e=0\) or
\(1\le e<m/4\). If \(w(S)\) is its bare incidence and local weight,
(8.21)--(8.25) give the nonnegative bound

\[
 \sum_{S\in\mathcal S(\mathbf L)}w(S)
  \le W(\mathbf L)\prod_{c\in\operatorname{cells}(\mathbf L)}
   \left(1+\sum_{1\le e<m_c/4}n^eR_{m_c,d_c}(e)\right).
                                                               \tag{8.25a}
\]

Thus a near-containment decoration is simply an endpoint cell of smaller
size \(m\) whose multiplicity has been lowered to \(m-e\) by fewer than
\(m/4\) stubs; the terminology records its geometric proximity to full
containment.

Distinguishing and then forgetting identical typed cells is exactly the
multinomial expansion, so there is no additional multiplicity. Since a
high skeleton is a matching and has at most \(k_{\mathrm{co}}\) cells, (8.25a) is
at most \(W(\mathbf L)\) times

\[
 (1+O((\ln n)^3/n))^{k_{\mathrm{co}}}=\exp(O((\ln n)^2)).                     \tag{8.26}
\]

\displayheading{Step III: middle cells with large residual mass}
It remains to cover the middle strip, rather than silently importing an
equitable-profile estimate. First expose a near-containment skeleton \(S\).
Its underlying set of high cells is a matching; write \(\mathcal M(S)\) for
that matching after multiplicities are forgotten, and let \(m_0\) be the
residual stub mass. Let \(\mathcal N(S)\) be the residual
event that there is no further near-containment cell. On \(\mathcal N(S)\),
every unexposed high cell is in the range
\(R_0<j\le3a/4+O(1)\). Define \(E_{\mathrm{mid}}(S)\) as the residual expectation of
the product of the local factors truncated to this middle range, multiplied
by \(\mathbf 1_{\mathcal N(S)}\). If \(\widetilde g\) denotes this
truncated factor, then on \(\mathcal N(S)\)

\[
 \widetilde g(r_{ab})
 \le 1+\sum_{R_0<j\le3a/4+O(1)}
          g(j)\mathbf1_{\{r_{ab}\ge j\}}.
\]

Expanding the product produces nonnegative monomials, each of which chooses
distinct residual cells and at most one threshold in each selected cell.
Apply Lemma~6.2 jointly to the entire threshold family before changing any
residual restriction. Only afterwards do we discard
\(\mathbf1_{\mathcal N(S)}\): this can only increase each nonnegative
monomial. The local factor remains truncated to the stated middle range, so
the relaxation does not create a new charge for a newly admitted
near-containment cell. If
\(m_0\ge n/(\ln n)^6\), expand over distinct residual cells and their threshold
demands. With
\(\theta_{ab}=\mathrm e\,d_ad'_b/m_0\le \mathrm e\,a^2/m_0\), Lemma 6.2 is applied jointly
before the constraints are dropped, and gives

\[
\begin{split}
 E_{\mathrm{mid}}(S)
 &\le\sum_{\substack{D\ \text{ distinct residual cells}\\
               R_0<j_{ab}\le3a/4+O(1)}}
       \prod_{ab\in D}g(j_{ab})
       \Prob{r_{ab}\ge j_{ab}\ \text{for every }ab\in D}\\
 &\le\prod_{ab}\left(1+
       \sum_{R_0<j\le3a/4+O(1)}
       g(j)\frac{\theta_{ab}^j}{j!}\right)
 \le\exp(\Xi_4),                                        \end{split}
\tag{8.26a}
\]

where

\[
 \Xi_4\le k_{\mathrm{co}}^2
 \sum_{a/2<j\le3a/4+O(1)}
  g(j)\frac{(\mathrm e\,a^2/m_0)^j}{j!}.                 \tag{8.27}
\]

Put \(\ell_n=\log_2 n\) and \(j=x\ell_n\). Uniformly for
\(1+o(1)\le x\le 3/2+o(1)\),

\[
\begin{split}
 \log_2\left[k_{\mathrm{co}}^2g(j)
       \frac{(\mathrm e\,a^2/m_0)^j}{j!}\right]
 &\le2\ell_n+\frac{j^2}{2}-j\log_2 m_0+O(j\log_2a)\\
 &\le-c\ell_n^2.                                        \end{split}
\tag{8.28}
\]

Indeed, the quadratic coefficient is
\(x^2/2-x\le -3/8+o(1)\)
on this interval. Hence
\(\Xi_4=2^{-\Omega(\ell_n^2)}\).

\displayheading{Step IV: middle cells with small residual mass}
If instead \(m_0<n/(\ln n)^6\), leave the entire residual matching unexposed.
For an upper bound, replace \(\mathbf1_{\mathcal N(S)}\) by \(1\) and the
truncated local product by the full product \(\prod_e g(r_e)\); both
replacements enlarge the nonnegative integrand. Every residual degree is at
most \(a\), so every residual cell has
\(r_e\le a\) and therefore \(\binom{r_e}{2}\le((a-1)/2)r_e\). The near
skeleton is a matching, so pointwise

\[
 \beta(\mathcal M(S)\cup H_{\mathrm{res}})
 \le|E(H_{\mathrm{res}})|\le m_0/2,
 \quad
 \sum_e\binom{r_e}{2}\le(a-1)m_0/2.                    \tag{8.29}
\]

All remaining local and topological factors together are therefore at
most \(\exp(O(am_0))=\exp(O(n/(\ln n)^5))\). Since \(g\ge1\) and
\(\beta\ge0\), summing the residual matching probability first gives the
explicit conditional bound

\[
 E_{\mathrm{mid}}(S)
 \le\Expres{
       \prod_e g(r_e)2^{\beta(\mathcal M(S)\cup H_{\mathrm{res}})}}
 \le\exp(Cam_0).                                         \tag{8.29a}
\]

\displayheading{Step V: assembly}
Combining (8.16), (8.25a)--(8.26a), and the two residual-mass branches
now gives the global inequality

\[
 \sum_{(\mathcal M,j)\ \mathrm{feasible}}w_{\mathrm{hi}}(\mathcal M,j)
  \le e^{O((\ln n)^2)}\left(\sum_{\mathbf L}W(\mathbf L)\right)
      \max\{e^{\Xi_4},e^{Cn/(\ln n)^5}\}.                     \tag{8.29b}
\]

Therefore

\[
 \sum_{(\mathcal M,j)\ \mathrm{feasible}}w_{\mathrm{hi}}(\mathcal M,j)
 \le\exp\{O(\sqrt{n{\ln n}})+O((\ln n)^2)+O(n/(\ln n)^5)\}
 =\exp\{o(n/(\ln n)^4)\}.                                   \tag{8.30}
\]

This is the skeleton-sum half of the factorization (9.2); no residual
attachment has yet been charged.
\end{proof}

\section{All residual local and cycle
attachments}\label{all-residual-local-and-cycle-attachments}

\displayheading{Contract for this section}
Section~8 supplied the sum of the high-skeleton weights
\(w_{\mathrm{hi}}\), but deliberately left the unexposed local and
cycle-space factors unpaid.  Here we condition on one feasible skeleton.
The remaining stubs form a finite bipartite configuration model with
prescribed residual degrees, and \(\mathcal A(\mathcal M,j)\) records exactly
its residual attachment.  Lemma~9.1 bounds this quantity uniformly;
Proposition~9.2 then multiplies that uniform bound by the skeleton sum (8.30).

\displayheading{Exact conditional decomposition}
Fix an actual canonical high skeleton \((\mathcal M,j)\) as in (8.2), and
let \(m_0=n-J\) be the residual total degree. Its residual degrees
are at most \(u_{\max}\). Define

\[
 \mathcal A(\mathcal M,j)=\Expres{
  \left(\prod_e g(r'_e)\right)
  2^{\beta(\mathcal M\cup H_{\mathrm{res}})}\mathbf1_{\mathcal E(\mathcal M,j)}
 },                                                      \tag{9.1}
\]

where \(\mathcal E(\mathcal M,j)\) imposes the residual cap
\(r'_e\le R_0\) and forbids a further
pair in a cell of \(\mathcal M\). By (6.4), (8.3), and the exact conditional
law, the normalized moment is exactly

\[
 \frac{\Exp{(Z_{\mathbf k}^{\mathrm{sgn}})^2}}
      {\Exp{Z_{\mathbf k}^{\mathrm{sgn}}}^2}
 =\sum_{(\mathcal M,j)}w_{\mathrm{hi}}(\mathcal M,j)
       \mathcal A(\mathcal M,j).                                \tag{9.2}
\]

The architecture is now visible in a single line: Section~8 bounds the sum
of \(w_{\mathrm{hi}}\), Lemma~9.1 bounds the supremum of \(\mathcal A\), and
Proposition~9.2 combines the two bounds.

The sum in (9.2) is over the finite set of feasible canonical high skeletons:
typed matchings \(\mathcal M\), allowed multiplicities \(j_e>R_0\), and the induced
nonnegative residual row and column degrees.

Here \(\Expres{\cdot}\) denotes expectation only over the unexposed matching of
the remaining stubs.  A \emph{residual attachment} is this conditional
contribution: the local factors \(g(r'_e)\) record residual cell
multiplicities, while \(2^{\beta(\mathcal M\cup H_{\mathrm{res}})}\)
records the associated cycle-space contribution after the high skeleton has
already been fixed.

\displayheading{Uniform asymptotic bound}
We next bound the exact conditional quantity (9.1) uniformly over its
finite family of feasible skeletons.

\phantomsection
\addcontentsline{toc}{subsection}{Lemma 9.1 (uniform attachment bound)}
\label{lemma-9.1-uniform-attachment-bound}
\begin{lemmabox}{Lemma 9.1 (uniform attachment bound)}

There is a deterministic sequence \(\varepsilon_n\to0\), independent of the
canonical high skeleton, such that for every feasible skeleton and all
sufficiently large \(n\),

\[
 \mathcal A(\mathcal M,j)\le
 \exp\!\left\{\varepsilon_n\frac n{(\ln n)^4}\right\}.       \tag{9.3}
\]

More precisely, for an absolute \(C\), it is at most \(\exp(C(\ln n)^8)\) when
\(m_0\ge n/(\ln n)^6\) and at most \(\exp(Cn/(\ln n)^5)\) otherwise. No lower bound
on \(\mathcal A\) is asserted or needed.

\end{lemmabox}

\begin{proof}\label{proof-large-residual-degree}
\displayheading{Large residual mass}

There are two possible ways a residual cell can matter.  It may contribute a
large local multiplicity factor \(g(r'_e)\), or it may be selected as part of
an even residual subgraph contributing to the cycle-space term.  The expansion
below assigns these two costs to different weights.  The quantities
\(\lambda^{\mathrm{loc}}_{ab}\) pay for multiplicity increments \(r'_e\ge3\),
whereas \(q_{ab}\) pays for a residual edge used by the even-subgraph sum.

Suppose \(m_0\ge n/(\ln n)^6\). Since

\[
 u_{\max}=(2+o(1))\log_2 n,\qquad
 \log_2 m_0=\log_2 n-O(\log_2\ln n),                           \tag{9.4}
\]

we have \(u_{\max}<3 \log_2 m_0\) for all sufficiently large
\(n\). For a residual cell outside \(\mathcal M\), put

\[
 \theta_{ab}=\frac{\mathrm e\, d_ad'_b}{m_0},\quad
 \Delta_x=g(x)-g(x-1),                                  \tag{9.5}
\]

and

\[
 \lambda^{\mathrm{loc}}_{ab}=\sum_{x=3}^{R_0}\Delta_x
                    \frac{\theta_{ab}^x}{x!},\qquad
 q_{ab}=\frac{\theta_{ab}^2}{2}+\lambda^{\mathrm{loc}}_{ab}.
                                                               \tag{9.6}
\]

Set both quantities to zero on \(\mathcal M\). Let
\(\theta_*=\mathrm e\,u_{\max}^2/m_0\). The ratio of consecutive majorants
\(g(x)\theta_*^x/x!\) is

\[
 \frac{2^x\theta_*}{x+1},                                \tag{9.7}
\]

and increases with \(x\). Hence the sequence is log-convex. Its
decreasing head is dominated by the \(x=3\) term. Consecutive
values of (9.7) grow by the factor
\(2(x+1)/(x+2)\ge 8/5\), so only \(O(1)\)
transition terms can have ratio between \(1/2\) and \(1\). Its
increasing tail is dominated by its right endpoint, and

\[
 \log_2\left(g(R_0)\frac{\theta_*^{R_0}}{R_0!}\right)
 \le\frac{R_0^2}{2}-R_0\log_2 m_0+O(R_0\log_2u_{\max})
 =-\Omega(R_0\log n),                                   \tag{9.8}
\]

using \(R_0=u_{\max}/2\) and (9.4). Thus, uniformly in the cell,

\[
 \lambda^{\mathrm{loc}}_{ab}\le C\theta_{ab}^3,\qquad
 q_{ab}\le C\theta_{ab}^2.                              \tag{9.9}
\]

Expand each capped local factor as

\[
 g(r'_e)=1+\sum_{x=3}^{R_0}\Delta_x
                   \mathbf1_{\{r'_e\ge x\}}.            \tag{9.10}
\]

If an edge is selected in a fixed even subgraph, use instead

\[
 \mathbf1_{\{r'_e\ge2\}}g(r'_e)
 =\mathbf1_{\{r'_e\ge2\}}
  +\sum_{x=3}^{R_0}\Delta_x\mathbf1_{\{r'_e\ge x\}}.   \tag{9.11}
\]

In the product expansion, a selected cycle-space edge contributes either
the base demand \(\mathbf1_{\{r'_e\ge2\}}\) or one increment term
\(\Delta_x\mathbf1_{\{r'_e\ge x\}}\), but not both in the same monomial.
Thus a cell of multiplicity at least three is not charged simultaneously as
a separate double-cell demand and as a higher-threshold local-factor demand.
Let \(R\) be the finite potential residual support graph consisting of all
residual row--column cells, and let
\(\mathcal C_{\rm even}(\mathcal M,R)\) be the finite family of edge sets
\(F\subseteq E(\mathcal M\cup R)\) having even degree at every vertex.  The
edges of \(\mathcal M\) have weight one below.

For clarity, the derivation of the next inequality has four steps.  First,
expand \(2^\beta\) by the cycle-space identity (6.7).  Second, use (9.10) on
edges outside the chosen even set and (9.11) on its selected residual edges.
Third, apply Lemma~6.2 jointly to every resulting threshold demand before
removing the cap or the no-return restriction.  Fourth, factor the remaining
local increments cell by cell.  Every coefficient and event indicator is
nonnegative, so each relaxation preserves the upper-bound direction.  This
gives

\[
 \mathcal A(\mathcal M,j)
  \le e^{\Lambda_{\mathrm{loc}}}
  \sum_{F\in\mathcal C_{\rm even}(\mathcal M,R)}
       \prod_{e\in F\setminus \mathcal M}q_e,\qquad
  \Lambda_{\mathrm{loc}}=\sum_e\lambda^{\mathrm{loc}}_e.
                                                               \tag{9.12}
\]

The right side now has a clean division of labour: the exponential term pays
for higher local multiplicities, while the weighted even-subgraph sum pays
for the cycle space. The remaining proof bounds these two quantities
independently.

By (9.9),

\[
 \sum_{a,b}\theta_{ab}^3
 =\frac{\mathrm e^3}{m_0^3}
   \left(\sum_a d_a^3\right)\left(\sum_b(d'_b)^3\right).
\]

Every residual degree is at most \(u_{\max}\), and both degree sums equal
\(m_0\).  Therefore
\(\sum_a d_a^3\le u_{\max}^2m_0\) and
\(\sum_b(d'_b)^3\le u_{\max}^2m_0\), which gives

\[
 \Lambda_{\mathrm{loc}}\le\frac C{m_0^3}
   \left(\sum_a d_a^3\right)\left(\sum_b(d'_b)^3\right)
 \le C\frac{u_{\max}^4}{m_0}.                                  \tag{9.13}
\]

For a fixed row \(a\), the quadratic bound in (9.9) also gives

\[
 \sum_b\theta_{ab}^2
 =\frac{\mathrm e^2d_a^2}{m_0^2}\sum_b(d'_b)^2
 \le C\frac{u_{\max}^3}{m_0},
\]

and the same calculation holds for every column.  Hence

\[
 \max_a\sum_bq_{ab}\ \vee\ \max_b\sum_aq_{ab}
 \le C\frac{u_{\max}^3}{m_0}=:\tau=o(1).                       \tag{9.14}
\]

\displayheading{Cycle-kernel estimate}
Every even edge set has a deterministic decomposition into edge-disjoint
simple cycles, and the map from an even edge set to that chosen cycle family
is injective because the original edge set is recovered by taking the
union. We now enlarge the image of this map to all finite families of simple
cycles, allowing cycles in the family to share edges. Since all cycle
weights are nonnegative, this relaxation can only increase the sum.  With
\(w(C)=\prod_{e\in C\setminus \mathcal M}q_e\) and
\(w(F)=\prod_{e\in F\setminus \mathcal M}q_e\), summing over all cycle
families and using \(1+x\le e^x\) gives

\[
 \sum_{F\text{ even}}\prod_{e\in F\setminus \mathcal M}q_e
 \le\exp\left\{\sum_{C\text{ simple cycle}}
                  \prod_{e\in C\setminus \mathcal M}q_e\right\}. \tag{9.15}
\]

Let \(K\) be the symmetric nonnegative kernel on the disjoint union of the
residual row and column vertices, with
\(K(a,b)=K(b,a)=q_{ab}\) on residual cells and zero elsewhere.
Equation (9.14) says that its maximum row sum is at most \(\tau\).

For cycles disjoint from \(\mathcal M\), orient the cycle and distinguish
one row vertex. A cycle of length \(2r\), \(r\ge2\), is thereby encoded by
one of the alternating \(K\)-walks of length \(2r\) from that vertex back to
itself. Dropping simplicity and the final return constraint only enlarges
the family. There are at most \(n\) possible distinguished row vertices,
and successive row-sum bounds give \(n\sum_{r\ge2}\tau^{2r}\). Hence

\[
 \sum_{C:C\cap \mathcal M=\varnothing}\prod_{e\in C}q_e
 \le\frac{n\tau^4}{1-\tau^2}.                            \tag{9.16}
\]

For the mixed cycles, define the positive residual-walk kernel

\[
 S=\sum_{\ell\ge1}K^\ell,\qquad
 \|S\|_\infty\le
 b:=\sum_{\ell\ge1}\tau^\ell=\frac\tau{1-\tau}.          \tag{9.17}
\]

This series is finite entrywise for all sufficiently large \(n\), since
\(\tau=o(1)\), and its entries are the total weights of positive residual
walks with prescribed endpoints. Let \(P\) be the symmetric adjacency
kernel of the matching \(\mathcal M\). It is a partial permutation, so
\(\|P\|_\infty\le1\).

Bound the mixed-cycle sum by first marking and orienting one of its matching
edges; this only overcounts. Across all cycles there are at most \(2h\)
possible oriented starts. If a marked cycle uses exactly \(r\ge1\) matching
edges, cutting it at those edges gives \(r\) nonempty
residual path segments: consecutive matching edges cannot occur because
\(\mathcal M\) is a matching. After the distinguished edge, the ordered
segment data injectively records the marked cycle. If simplicity, distinctness of
the later matching edges, and the final endpoint constraint are discarded,
the total remaining weight for the fixed oriented start is at most
\[
 \bigl\|S(PS)^{r-1}\bigr\|_\infty\le b^r.
\]
The initial choice is the only factor depending on \(h\); every later
matching traversal is already represented by \(P\). Summing over the
oriented initial edge and over \(r\) gives

\[
 \sum_{C:C\cap \mathcal M\ne\varnothing}
       \prod_{e\in C\setminus \mathcal M}q_e
 \le2h\sum_{r\ge1}b^r\le Ch\tau.                        \tag{9.18}
\]

Every high cell uses more than \(u_{\max}/2\) stubs, so
\(h<2n/u_{\max}\).  At the cutoff \(m_0\ge n/(\ln n)^6\),

\[
 \tau=O((\ln n)^9/n),\qquad
 h\tau=O((\ln n)^8),\qquad
 \frac{u_{\max}^4}{m_0}+n\tau^4=o(1).
\]

Combining (9.12)--(9.18),

\[
 \mathcal A(\mathcal M,j)
 \le\exp\left[C\left\{\frac{u_{\max}^4}{m_0}+n\tau^4+h\tau\right\}\right]
 \le\exp(C'(\ln n)^8)                                          \tag{9.19}
\]

at the cutoff \(m_0\ge n/(\ln n)^6\).

\displayheading{Small residual mass}\label{proof-small-residual-degree}

Suppose \(m_0<n/(\ln n)^6\). Since \(\mathcal M\) is a
matching, it is a forest, and adding a residual support edge increases
cycle rank by at most one. Hence

\[
 \beta(\mathcal M\cup H_{\mathrm{res}})\le|E(H_{\mathrm{res}})|\le m_0/2.           \tag{9.20}
\]

Also, pointwise,

\[
 \sum_e\binom{r'_e}{2}
 \le\frac{u_{\max}-1}{2}\sum_er'_e=\frac{u_{\max}-1}{2}m_0.           \tag{9.21}
\]

Equations (6.5), (9.20), and (9.21) bound the integrand in (9.1) by
\(2^{u_{\max}m_0/2}\). Therefore

\[
 \mathcal A(\mathcal M,j)
 \le\exp(O(u_{\max}m_0))\le\exp(Cn/(\ln n)^5).             \tag{9.22}
\]

Both exponents in (9.19) and (9.22) are \(o(n/(\ln n)^4)\).
\end{proof}

\phantomsection
\addcontentsline{toc}{subsection}{Proposition 9.2 (normalized signed second moment)}
\label{proposition-9.2-normalized-signed-second-moment}
\begin{propositionbox}{Proposition 9.2 (normalized signed second moment)}

For the exact midpoint profile,

\[
 \Lambda_n:=\ln\frac{\Exp{(Z_{\mathbf k}^{\mathrm{sgn}})^2}}
                      {\Exp{Z_{\mathbf k}^{\mathrm{sgn}}}^2}
 =o(n/(\ln n)^4).                                               \tag{9.23}
\]

\end{propositionbox}

\begin{proof}\label{proof-8}

Equation (9.2) is an exact canonical decomposition.  Lemma 8.3 sums all
bare high skeletons by \(\exp(o(n/(\ln n)^4))\), including the empty
skeleton, all dense endpoint tables, every near-containment decoration,
and the two middle-strip regimes. Lemma 9.1 bounds the conditional
residual factor by the same exponential scale uniformly over every
skeleton.  Thus (9.2) gives the explicit estimate

\[
 \frac{\Exp{(Z_{\mathbf k}^{\mathrm{sgn}})^2}}
      {\Exp{Z_{\mathbf k}^{\mathrm{sgn}}}^2}
 \le
 \left(\sup_{(\mathcal M,j)}\mathcal A(\mathcal M,j)\right)
 \sum_{(\mathcal M,j)}w_{\mathrm{hi}}(\mathcal M,j)
 =\exp\{o(n/(\ln n)^4)\}.
\]

This proves the upper bound in (9.23). The
normalized second moment is at least one by variance, so its logarithm
is nonnegative and (9.23) follows.
\end{proof}

\section{Rare-event amplification}\label{rare-event-amplification}

Proposition 9.2 and (1.4) give the seed

\[
 \Prob{Z_{\mathbf k}^{\mathrm{sgn}}>0}\ge e^{-\Lambda_n}.       \tag{10.1}
\]

This is only a possibly rare seed.  The purpose of this section is to turn
it into a high-probability cocoloring event while adding fewer parts than the
root separation can absorb.

The proof of \citet[Theorem~1]{heckel-2025-difference} also begins with a
rare second-moment seed and then uses vertex-exposure concentration of
\(\zeta(G_n)\) to obtain a high-probability statement.  We use the same
seed-to-typical strategic principle, but not that theorem as a black box:
Lemma 10.2 proves the quantitative implication needed here for an arbitrary
seed exponent \(\Lambda_n\), and Lemma 10.1 supplies the simultaneous
leftover coloring that controls the added parts.

The ordinary-coloring concentration argument motivating this amplification
appears in \citet[Theorem~1]{scott-2008-2017}. Lemmas 10.1 and 10.2 prove the precise
simultaneous-leftover and rare-seed forms required here.

By the observation after (5.19), this implies

\[
 \Prob{\zeta(G_n)\le k_{\mathrm{co}}}\ge e^{-\Lambda_n}.        \tag{10.2}
\]

We now prove the amplification needed to turn this possibly rare event
into a typical one.

\phantomsection
\addcontentsline{toc}{subsection}{Lemma 10.1 (simultaneous leftover coloring)}
\label{lemma-10.1-simultaneous-leftover-coloring}
\begin{lemmabox}{Lemma 10.1 (simultaneous leftover coloring)}

There is an absolute \(C\) such that, with probability \(1-o(1)\), every
\(S\subseteq[n]\) satisfies

\[
 \chi(G_n[S])\le C\frac{|S|}{{\ln n}}+n^{1/3}.                 \tag{10.3}
\]

\end{lemmabox}

\paragraph{Proof.}\label{proof-9}

Let \(H\) be the complement of \(G_n\). Put
\(u_0=\lceil n^{1/4}\rceil\). For any fixed \(u_0\)-set, (1.5)
gives probability
\(\exp(-\Omega(u_0^2))\)
that its \(H\)-edge density is below \(1/4\). There are at
most \(\binom n{u_0}\) such sets, and for an absolute \(c>0\),

\[
 \binom n{u_0}e^{-cu_0^2}
 \le\exp\{u_0\ln(\mathrm e n/u_0)-cu_0^2\}=o(1).
\]

Thus a union bound shows that, with
probability \(1-o(1)\), every \(u_0\)-set has density at
least \(1/4\). By averaging over its \(u_0\)-subsets, the
same is then true of every larger set.

Start with any vertex set of size at least \(n^{1/3}\). While
its current size is at least \(u_0\), choose a vertex of maximum
\(H\)-degree and replace the current set by its
\(H\)-neighborhood there. Density at least \(1/4\) implies
that this neighborhood has size at least one quarter of the current
size, up to an additive one. If \(s_t\) is the current size after \(t\)
choices, then

\[
 s_{t+1}\ge(s_t-1)/4,\qquad s_t\ge4^{-t}s_0-1/3.         \tag{10.3a}
\]

The chosen vertices form a clique in \(H\): after a vertex is chosen, the
current set is replaced by its \(H\)-neighbourhood, so every later choice is
adjacent in \(H\) to every earlier one. Starting from any
\(\lceil n^{1/3}\rceil\)-subset, (10.3a) stays above \(n^{1/4}\) for at
least \(\lfloor {\ln n}/(13\ln4)\rfloor\) steps. Thus the procedure constructs
at least \(c{\ln n}\) vertices for an absolute \(c>0\), and these vertices form
an independent set in \(G_n\).

For an arbitrary \(S\), repeatedly remove such independent sets and
give each a new color until fewer than \(n^{1/3}\) vertices
remain; color the rest singly.  Every removed set has at least
\(c\ln n\) vertices, so the total number of colors is at most

\[
 \left\lceil\frac{|S|}{c\ln n}\right\rceil+n^{1/3}.
\]

After enlarging the absolute constant, this is (10.3), simultaneously for
every \(S\). \(\square\)

\phantomsection
\addcontentsline{toc}{subsection}{Lemma 10.2 (amplification from a seed)}
\label{lemma-10.2-amplification-from-a-seed}
\begin{lemmabox}{Lemma 10.2 (amplification from a seed)}

Suppose deterministic \(k_n,\Lambda_n\ge0\) satisfy

\[
 \Prob{\zeta(G_n)\le k_n}\ge e^{-\Lambda_n}.             \tag{10.4}
\]

There is a deterministic \(\varepsilon_n=o(1)\), independent of
\(k_n,\Lambda_n,r\), such that, uniformly for every deterministic choice
\(r=r(n)>0\),

\[
\begin{split}
 \mathbb{P}\!\Bigg(\zeta(G_n)>k_n+C\bigg(
 &\frac{\sqrt{n\Lambda_n}+\sqrt{nr}}{\ln n}
   +n^{1/3}+1\bigg)\Bigg)\\
 &\le e^{-r}+\varepsilon_n.                              \end{split}
\tag{10.5}
\]

\end{lemmabox}

\paragraph{Proof.}\label{proof-10}

For an induced set \(W\), let
\(\zeta(W)=\zeta(G_n[W])\),
and define

\[
 S_k=\max\{|W|:\zeta(W)\le k_n\}.                       \tag{10.6}
\]

Expose the random graph in \(n-1\) independent vertex blocks, where
block \(v\) contains the edges from \(v\) to earlier vertices.
Changing one block changes \(S_k\) by at most one. Indeed, after deleting
the affected vertex, every feasible induced set loses at most one vertex,
and the two graph configurations agree on all remaining edges. Thus a
maximizer in either configuration yields a feasible set of size at least one
less in the other configuration. Therefore (1.3) applies.

Since \(S_k=n\) exactly when \(\zeta(G_n)\le k_n\),
(10.4) and the upper one-sided bounded-differences tail give

\[
 e^{-\Lambda_n}
 \le\Prob{S_k-\Exp{S_k}\ge n-\Exp{S_k}}
 \le\exp\!\left\{-\frac{2(n-\Exp{S_k})^2}{n-1}\right\}.
\]

Taking logarithms and rearranging imply

\[
 n-\Exp{S_k}\le\sqrt{(n-1)\Lambda_n/2}.                  \tag{10.7}
\]

The lower tail with radius \(\sqrt{(n-1)r/2}\) gives, except with
probability \(e^{-r}\),

\[
 n-S_k\le
 \sqrt{(n-1)\Lambda_n/2}+\sqrt{(n-1)r/2}.                \tag{10.8}
\]

Choose a maximizing set \(W\) and put
\(V_{\mathrm{left}}=[n]\setminus W\). Concatenating a
\(k_n\)-part cocoloring of \(G[W]\) with an ordinary
coloring of \(G[V_{\mathrm{left}}]\) gives

\[
 \zeta(G_n)\le k_n+\chi(G_n[V_{\mathrm{left}}]).         \tag{10.9}
\]

Intersect (10.8) with the simultaneous event in Lemma 10.1 and apply
(10.3).  Since \(|V_{\mathrm{left}}|=n-S_k\), equations (10.3), (10.8), and (10.9)
give exactly the number of additional parts displayed in (10.5).
The simultaneous event's failure probability is the same deterministic
\(\varepsilon_n=o(1)\) for every choice of the three deterministic
parameters. This proves (10.5). \(\square\)

Apply Lemma 10.2 to (10.2). Put

\[
 B_n=\frac n{(\ln n)^4},\qquad r_n=\sqrt{B_n}.                 \tag{10.10}
\]

Then \(r_n\to\infty\), and Proposition 9.2 gives

\[
 \frac{\sqrt{n\Lambda_n}}{\ln n}=o(n/(\ln n)^3),\qquad
 \frac{\sqrt{nr_n}}{\ln n}=o(n/(\ln n)^3),\qquad
 n^{1/3}=o(n/(\ln n)^3).                                       \tag{10.11}
\]

Define the deterministic sequence

\[
 a_n=C\left(
 \frac{\sqrt{n\Lambda_n}+\sqrt{nr_n}}{\ln n}+n^{1/3}+1\right).
                                                               \tag{10.12}
\]

Then (10.11) and Lemma 10.2 give

\[
 a_n=o(n/(\ln n)^3),\qquad
 \Prob{\zeta(G_n)>k_{\mathrm{co}}+a_n}
 \le e^{-r_n}+\varepsilon_n\longrightarrow0.            \tag{10.13}
\]

\section{Completion of the proof}\label{completion-of-the-proof}

Intersect \(\{\chi(G_n)>k_\chi^-\}\), whose probability tends to one by
(4.5), with \(\{\zeta(G_n)\le k_{\mathrm{co}}+a_n\}\), whose probability tends to
one by (10.13). No independence is required; a union bound suffices.
Equation (5.20) then gives, with high probability,

\[
 \chi(G_n)-\zeta(G_n)
 \ge\left(\frac{(\ln 2)^2\gamma_4}{16}-o(1)\right)
        \frac n{(\ln n)^3}.                                    \tag{11.1}
\]

For every sufficiently large \(n\), the right side is at least

\[
 c_*\frac n{(\ln n)^3},\qquad
 c_*:=\frac{(\ln 2)^2\gamma_4}{32}
 =\frac{(\ln 2)^2}{32}\ln\frac{200}{153}>0.              \tag{11.2}
\]

This proves (0.1). Since \(n/(\ln n)^3\) tends to infinity, for
every fixed \(K\),

\[
 \Prob{\chi(G_n)-\zeta(G_n)\ge K}\longrightarrow1.      \tag{11.3}
\]

Every estimate used above is uniform for \(\delta \in [0,1)\) and for the
exact tangent-rounded finite-\(n\) profile. The conclusion therefore holds
along all sufficiently large integers, including sequences approaching
either side of every independence-number jump. The phase has not been
averaged away or discarded: the result is a full-sequence high-probability
statement, not a density-subset or subsequence statement. \(\blacksquare\)

\section*{Reproducibility and formalisation}

The complete proof sources and reproducibility material are collected in the
numbered dossiers of the public
\href{https://github.com/SamPetkov/Erdos}{\texttt{Erdos} repository}; the
present paper is in \texttt{625/}. All theorem statements and proof text are
available with the manuscript sources. The dossier includes the TeX source,
bibliography, reproducible build material, audit records, and the current
partial Lean~4 formalisation.

The \texttt{625/verification/} directory additionally contains an
independently written Python checker, its reproduced output, and a Markdown
verification report. These files provide supplementary finite consistency
checks and an auditable record of what was tested; they are not a substitute
for external peer review or for a complete formal proof.

The Lean checkpoint, developed using ChatGPT-5.6 and Aristotle
\citep{achim-et-al-2025}, is provided as a self-contained source file. Each
formalised statement is decomposed into smaller claims, with its hypotheses
and conclusion checked explicitly. This is a partial machine-checked
formalisation, not a claim that the entire paper has been formalised.

\section*{AI assistance}

The proof and formalisation were completed with extensive assistance from
ChatGPT-5.6 Sol Ultra and Aristotle. Separate adversarial verification passes
were run with ChatGPT-5.6 Sol Pro; these passes are not external peer review.
In particular, ChatGPT-5.6 proposed the four-size signed profile with class sizes
\(\alpha-2,\alpha-3,\alpha-4,\alpha-5\), which led to the phase-uniform
first-moment separation used in Section~5. The author reviewed every
incorporated suggestion and assumes full responsibility for the arguments,
citations, and final manuscript.

\section*{Funding}

The author received no funding for this work.

\section*{Competing interests}

The author declares no competing interests.

\renewcommand{\bibsection}{\section*{References}}
\bibliographystyle{unsrtnat}
\bibliography{references}

\end{document}
